### 2021-12-15 paper.tex
Add LaTeX files

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage[review]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{Final Paper for LING 380}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{Theo Sandstrom \\
  \texttt{theo.sandstrom@yale.edu} \\\And
  Trey Skidmore \\
  \texttt{trey.skidmore@yale.edu} \\\And
  Prastik Mohanraj \\
  \texttt{prastik.mohanraj@yale.edu}}

\begin{document}
\maketitle
\begin{abstract}
This document is a supplement to the general instructions for *ACL authors. It contains instructions for using the \LaTeX{} style files for ACL conferences. 
The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
These instructions should be used both for papers submitted for review and for final versions of accepted papers.
\end{abstract}

\section{Introduction}

These instructions are for authors submitting papers to *ACL conferences using \LaTeX. They are not self-contained. All authors must follow the general instructions for *ACL proceedings,\footnote{\url{http://acl-org.github.io/ACLPUB/formatting.html}} and this document contains additional instructions for the \LaTeX{} style files.

The templates include the \LaTeX{} source of this document (\texttt{acl.tex}),
the \LaTeX{} style file used to format it (\texttt{acl.sty}),
an ACL bibliography style (\texttt{acl\_natbib.bst}),
an example bibliography (\texttt{custom.bib}),
and the bibliography for the ACL Anthology (\texttt{anthology.bib}).

\section{Engines}

To produce a PDF file, pdf\LaTeX{} is strongly recommended (over original \LaTeX{} plus dvips+ps2pdf or dvipdf). Xe\LaTeX{} also produces PDF files, and is especially suitable for text in non-Latin scripts.

\section{Preamble}

The first line of the file must be
\begin{quote}
\begin{verbatim}
\documentclass[11pt]{article}
\end{verbatim}
\end{quote}

To load the style file in the review version:
\begin{quote}
\begin{verbatim}
\usepackage[review]{acl}
\end{verbatim}
\end{quote}
For the final version, omit the \verb|review| option:
\begin{quote}
\begin{verbatim}
\usepackage{acl}
\end{verbatim}
\end{quote}

To use Times Roman, put the following in the preamble:
\begin{quote}
\begin{verbatim}
\usepackage{times}
\end{verbatim}
\end{quote}
(Alternatives like txfonts or newtx are also acceptable.)

Please see the \LaTeX{} source of this document for comments on other packages that may be useful.

Set the title and author using \verb|\title| and \verb|\author|. Within the author list, format multiple authors using \verb|\and| and \verb|\And| and \verb|\AND|; please see the \LaTeX{} source for examples.

By default, the box containing the title and author names is set to the minimum of 5 cm. If you need more space, include the following in the preamble:
\begin{quote}
\begin{verbatim}
\setlength\titlebox{<dim>}
\end{verbatim}
\end{quote}
where \verb|<dim>| is replaced with a length. Do not set this length smaller than 5 cm.

\section{Document Body}

\subsection{Footnotes}

Footnotes are inserted with the \verb|\footnote| command.\footnote{This is a footnote.}

\subsection{Tables and figures}

See Table~\ref{tab:accents} for an example of a table and its caption.
\textbf{Do not override the default caption sizes.}

\begin{table}
\centering
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular}
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\
\hline
\end{tabular}
\caption{Example commands for accented characters, to be used in, \emph{e.g.}, Bib\TeX{} entries.}
\label{tab:accents}
\end{table}

\subsection{Hyperlinks}

Users of older versions of \LaTeX{} may encounter the following error during compilation: 
\begin{quote}
\tt\verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.
\end{quote}
This happens when pdf\LaTeX{} is used and a citation splits across a page boundary. The best way to fix this is to upgrade \LaTeX{} to 2018-12-01 or later.

\subsection{Citations}

\begin{table*}
\centering
\begin{tabular}{lll}
\hline
\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
\hline
\citep{Gusfield:97} & \verb|\citep| & \verb|\cite| \\
\citealp{Gusfield:97} & \verb|\citealp| & no equivalent \\
\citet{Gusfield:97} & \verb|\citet| & \verb|\newcite| \\
\citeyearpar{Gusfield:97} & \verb|\citeyearpar| & \verb|\shortcite| \\
\hline
\end{tabular}
\caption{\label{citation-guide}
Citation commands supported by the style file.
The style is based on the natbib package and supports all natbib citation commands.
It also supports commands defined in previous ACL style files for compatibility.
}
\end{table*}

Table~\ref{citation-guide} shows the syntax supported by the style files.
We encourage you to use the natbib styles.
You can use the command \verb|\citet| (cite in text) to get ``author (year)'' citations, like this citation to a paper by \citet{Gusfield:97}.
You can use the command \verb|\citep| (cite in parentheses) to get ``(author, year)'' citations \citep{Gusfield:97}.
You can use the command \verb|\citealp| (alternative cite without parentheses) to get ``author, year'' citations, which is useful for using citations within parentheses (e.g. \citealp{Gusfield:97}).

\subsection{References}

\nocite{Ando2005,borschinger-johnson-2011-particle,andrew2007scalable,rasooli-tetrault-2015,goodman-etal-2016-noise,harper-2014-learning}

The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
If your own bib file is named \texttt{custom.bib}, then placing the following before any appendices in your \LaTeX{} file will generate the references section for you:
\begin{quote}
\begin{verbatim}
\bibliographystyle{acl_natbib}
\bibliography{custom}
\end{verbatim}
\end{quote}

You can obtain the complete ACL Anthology as a Bib\TeX{} file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
To include both the Anthology and your own .bib file, use the following instead of the above.
\begin{quote}
\begin{verbatim}
\bibliographystyle{acl_natbib}
\bibliography{anthology,custom}
\end{verbatim}
\end{quote}

Please see Section~\ref{sec:bibtex} for information on preparing Bib\TeX{} files.

\subsection{Appendices}

Use \verb|\appendix| before any appendix section to switch the section numbering over to letters. See Appendix~\ref{sec:appendix} for an example.

\section{Bib\TeX{} Files}
\label{sec:bibtex}

Unicode cannot be used in Bib\TeX{} entries, and some ways of typing special characters can disrupt Bib\TeX's alphabetization. The recommended way of typing special characters is shown in Table~\ref{tab:accents}.

Please ensure that Bib\TeX{} records contain DOIs or URLs when possible, and for all the ACL materials that you reference.
Use the \verb|doi| field for DOIs and the \verb|url| field for URLs.
If a Bib\TeX{} entry has a URL or DOI field, the paper title in the references section will appear as a hyperlink to the paper, using the hyperref \LaTeX{} package.

\section*{Acknowledgements}

This document has been adapted
by Steven Bethard, Ryan Cotterell and Rui Yan
from the instructions for earlier ACL and NAACL proceedings, including those for 
ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
NAACL 2019 by Stephanie Lukin and Alla Roskovskaya, 
ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu, 
NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
ACL 2017 by Dan Gildea and Min-Yen Kan, 
NAACL 2017 by Margaret Mitchell, 
ACL 2012 by Maggie Li and Michael White, 
ACL 2010 by Jing-Shin Chang and Philipp Koehn, 
ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui, 
ACL 2005 by Hwee Tou Ng and Kemal Oflazer, 
ACL 2002 by Eugene Charniak and Dekang Lin, 
and earlier ACL and EACL formats written by several people, including
John Chen, Henry S. Thompson and Donald Walker.
Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.

% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}
\bibliographystyle{acl_natbib}

\appendix

\section{Example Appendix}
\label{sec:appendix}

This is an appendix.

\end{document}

### 2021-12-16 paper.tex
add surprisal section

@@
 % Remove the "review" option to generate the final version.
-\usepackage[review]{acl}
+\usepackage{acl}
 
@@
 \author{Theo Sandstrom \\
-  \texttt{theo.sandstrom@yale.edu} \\\And
+  \texttt{theo.sandstrom@yale.edu} \\
   Trey Skidmore \\
-  \texttt{trey.skidmore@yale.edu} \\\And
+  \texttt{trey.skidmore@yale.edu} \\
   Prastik Mohanraj \\
@@
 
-\section{Engines}
+\section{Methods}
+\subsection{Language Models}
+\subsection{Surprisal}
+We asses the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric that is defined in Wilcox et.
+al.
+(2018).
+The metric is defined to be the log inverse probability:
+\[S(x_i) = -log_2(p_i|h_{i-1}),\]
+where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is the softmax activation. The surprisal is measured in bits because we take a base two logarithm.
 
-To produce a PDF file, pdf\LaTeX{} is strongly recommended (over original \LaTeX{} plus dvips+ps2pdf or dvipdf).
-Xe\LaTeX{} also produces PDF files, and is especially suitable for text in non-Latin scripts.
+A high surprisal metric means that a given word was unexpected in its context.
+The metric is known to correlate directly with human sentence processing difficulty (Hale, 2001; Levy, 2008; Smith and Levy, 2013).
+In order to probe the model’s understanding of negative polarity licensing, we will construct a synthetic dataset which contained both grammatical examples, in which negative polarity items are correctly licensed, and ungrammatical examples, in which the licensor is either not present or is not in a c-command relationship with the licensee.
 

### 2021-12-16 src/main.py
Add surprisal tracking code

@@ -1,7 +1,7 @@
 import argparse
-import time
-import math
 import os
+import pickle
+import pandas as pd
 
 import torch
 import torch.nn as nn
@@ -9,217 +9,99 @@ import torch.nn as nn
 import data
 import model
 
-
-parser = argparse.ArgumentParser(description='LSTM Language Model')
-parser.add_argument('--data', type=str, default='./data/wikipedia',
-                    help='location of data corpus')
+parser = argparse.ArgumentParser(description='Evaluate language model on examples')
+parser.add_argument('--model', type=str, default='lstm.pt',
+                    help="location of model to evaluate")
+parser.add_argument('--evaluate', type=str, 
+                    help='location of example sentences to evaluate')
+parser.add_argument('--dictionary', type=str, default='dictionary.p',
+                    help='location of the dictionary to use')
 parser.add_argument('--embedding_size', type=int, default=200,
                     help='size of word embeddings')
-parser.add_argument('--hidden_size', type=int, default=200,
+parser.add_argument('--hidden_size', type=int, default=650,
                     help='number of hidden units per layer')
 parser.add_argument('--num_layers', type=int, default=2,
                     help='number of layers')
-parser.add_argument('--lr', type=float, default=20,
-                    help='initial learning rate')
-parser.add_argument('--clip', type=float, default=0.25,
-                    help='gradient clipping')
-parser.add_argument('--epochs', type=int, default=40,
-                    help='upper epoch limit')
-parser.add_argument('--batch_size', type=int, default=20, metavar='N',
-                    help='batch size')
-parser.add_argument('--bptt', type=int, default=35,
-                    help='sequence length')
 parser.add_argument('--dropout', type=float, default=0.2,
                     help='dropout applied to layers (0 = no dropout)')
 parser.add_argument('--tied', action='store_true',
                     help='tie the word embedding and softmax weights')
-parser.add_argument('--seed', type=int, default=1111,
-                    help='random seed')
-parser.add_argument('--cuda', action='store_true',
-                    help='use CUDA')
-parser.add_argument('--log-interval', type=int, default=200, metavar='N',
-                    help='report interval')
-parser.add_argument('--save', type=str, default='model.pt',
-                    help='path to save the final model')
[…]

### 2021-12-16 src/model.py

@@ -16,20 +16,6 @@ class RNNModel(nn.Module):
                            dropout=dropout)
         self.decoder = nn.Linear(hidden_size, vocab_size)
 
-        # Optionally tie weights as in:
-        # "Using the Output Embedding to Improve Language Models" (Press & Wolf
-        # 2016)
-        # https://arxiv.org/abs/1608.05859
-        # and
-        # "Tying Word Vectors and Word Classifiers: A Loss Framework for
-        # Language Modeling" (Inan et al. 2016)
-        # https://arxiv.org/abs/1611.01462
-        if tie_weights:
-            if hidden_size != input_size:
-                raise ValueError('When using the tied flag, hidden_size must \
-                be equal to emsize')
-            self.decoder.weight = self.encoder.weight
-
         self.init_weights()
 
         self.hidden_size = hidden_size
@@ -46,7 +32,6 @@ class RNNModel(nn.Module):
         output, hidden = self.rnn(emb, hidden)
         output = self.drop(output)
         decoded = self.decoder(output)
-        decoded = decoded.view(-1, self.vocab_size)
         return F.log_softmax(decoded, dim=1), hidden
 
     def init_hidden(self, bsz):

### 2021-12-16 src/train.py

@@ -1,0 +1,225 @@
+import argparse
+import time
+import math
+import os
+
+import torch
+import torch.nn as nn
+
+import data
+import model
+
+
+parser = argparse.ArgumentParser(description='LSTM Language Model')
+parser.add_argument('--data', type=str, default='./data/wikipedia',
+                    help='location of data corpus')
+parser.add_argument('--embedding_size', type=int, default=200,
+                    help='size of word embeddings')
+parser.add_argument('--hidden_size', type=int, default=200,
+                    help='number of hidden units per layer')
+parser.add_argument('--num_layers', type=int, default=2,
+                    help='number of layers')
+parser.add_argument('--lr', type=float, default=20,
+                    help='initial learning rate')
+parser.add_argument('--clip', type=float, default=0.25,
+                    help='gradient clipping')
+parser.add_argument('--epochs', type=int, default=40,
+                    help='upper epoch limit')
+parser.add_argument('--batch_size', type=int, default=20, metavar='N',
+                    help='batch size')
+parser.add_argument('--bptt', type=int, default=35,
+                    help='sequence length')
+parser.add_argument('--dropout', type=float, default=0.2,
+                    help='dropout applied to layers (0 = no dropout)')
+parser.add_argument('--tied', action='store_true',
+                    help='tie the word embedding and softmax weights')
+parser.add_argument('--seed', type=int, default=1111,
+                    help='random seed')
+parser.add_argument('--cuda', action='store_true',
+                    help='use CUDA')
+parser.add_argument('--log-interval', type=int, default=200, metavar='N',
+                    help='report interval')
+parser.add_argument('--save', type=str, default='model.pt',
+                    help='path to save the final model')
+parser.add_argument('--dry-run', action='store_true',
+                    help='verify the code and the model')
+
+args = parser.parse_args()
+
+
+# Set the random seed manually for reproducibility.
+torch.manual_seed(args.seed)
+if torch.cuda.is_available():
+    if not args.cuda:
+        print("WARNING: You have a CUDA device, so you should probably run \
+        with --cuda")
+
+device = torch.device("cuda" if args.cuda else "cpu")
+
+###############################################################################
+# Load data
+###############################################################################
+
[…]

### 2021-12-17 src/main.py
Fix bug with loading model onto GPU

@@ -2,6 +2,7 @@ import argparse
 import os
 import pickle
 import pandas as pd
+import matplotlib.pyplot as plt
 
 import torch
 import torch.nn as nn
@@ -36,7 +37,7 @@ model = model.RNNModel(len(dictionary), args.embedding_size, args.hidden_size,
                        args.num_layers, args.dropout, args.tied)
 
 with open(args.model, 'rb') as f:
-    model = torch.load(f)
+    model = torch.load(f).to(torch.device("cpu"))
     model.rnn.flatten_parameters()
 
 def read_examples_file_as_tensor(examples, dictionary):
@@ -100,8 +101,9 @@ def plot_surprisal(vocab, model, batch):
         
         series = pd.Series(seq_surprisal, seq_words)
         series.plot()
+        plt.show()
 
 
 batch = read_examples_file_as_tensor(args.evaluate, dictionary)
 plot_surprisal(dictionary, model, batch)
-plt.show()
+

### 2021-12-17 src/main.py
Fix bug

@@ -39,6 +39,7 @@ model = model.RNNModel(len(dictionary), args.embedding_size, args.hidden_size,
 with open(args.model, 'rb') as f:
     model = torch.load(f).to(torch.device("cpu"))
     model.rnn.flatten_parameters()
+    model.eval()
 
 def read_examples_file_as_tensor(examples, dictionary):
     """Reads a text file with one example sentence on each line, and returns a
@@ -68,26 +69,27 @@ def surprisal(model, batch):
     """
     initial_hidden = torch.zeros(args.num_layers, batch.size(0),
                                  args.hidden_size)
-    model_output, _ = model(batch.transpose(0, 1),
-                            (initial_hidden,
-                             initial_hidden))
-    model_output = model_output.transpose(0, 1)
+    with torch.no_grad():
+        model_output, _ = model(batch.transpose(0, 1),
+                                (initial_hidden,
+                                 initial_hidden))
+        model_output = model_output.transpose(0, 1)
 
-    # Throw away the first column, because it has no preceding context.
-    next_word = torch.narrow(batch, 1, 1, batch.size(1) - 1)
-    # Throw away the last column because we don't know what should come next.
-    model_output = torch.narrow(model_output, 1, 0, batch.size(1) - 1)
-
-    # Squeeze and unsqueeze to make dimensions line up
-    next_word = next_word.unsqueeze(-1)
-    next_word_log_prob = torch.gather(model_output, -1, next_word)
-    next_word_log_prob = next_word_log_prob.squeeze()
-
-    # The model returns the natural log of the probability, so we divide by
-    # log(2) to get the log base 2.
-    surprisal = -next_word_log_prob / torch.log(torch.tensor(2))
-    # Set the surprisal of the initial word to be zero
-    surprisal = torch.cat([torch.zeros(surprisal.size(0), 1), surprisal], 1)
+        # Throw away the first column, because it has no preceding context.
+        next_word = torch.narrow(batch, 1, 1, batch.size(1) - 1)
+        # Throw away the last column because we don't know what should come next.
+        model_output = torch.narrow(model_output, 1, 0, batch.size(1) - 1)
+        
+        # Squeeze and unsqueeze to make dimensions line up
+        next_word = next_word.unsqueeze(-1)
+        next_word_log_prob = torch.gather(model_output, -1, next_word)
+        next_word_log_prob = next_word_log_prob.squeeze()
+        
[…]

### 2021-12-17 src/train.py
Update default parameters in train.py

@@ -11,11 +11,11 @@ import model
 
 
 parser = argparse.ArgumentParser(description='LSTM Language Model')
-parser.add_argument('--data', type=str, default='./data/wikipedia',
+parser.add_argument('--data', type=str, default='./data/wikitext-2',
                     help='location of data corpus')
 parser.add_argument('--embedding_size', type=int, default=200,
                     help='size of word embeddings')
-parser.add_argument('--hidden_size', type=int, default=200,
+parser.add_argument('--hidden_size', type=int, default=650,
                     help='number of hidden units per layer')
 parser.add_argument('--num_layers', type=int, default=2,
                     help='number of layers')
@@ -25,7 +25,7 @@ parser.add_argument('--clip', type=float, default=0.25,
                     help='gradient clipping')
 parser.add_argument('--epochs', type=int, default=40,
                     help='upper epoch limit')
-parser.add_argument('--batch_size', type=int, default=20, metavar='N',
+parser.add_argument('--batch_size', type=int, default=128, metavar='N',
                     help='batch size')
 parser.add_argument('--bptt', type=int, default=35,
                     help='sequence length')

### 2021-12-17 paper.tex
Add intro to analysis sections

@@
 \usepackage{acl}
-
+\usepackage{gb4e}
 % Standard package includes
@@
 \begin{abstract}
-This document is a supplement to the general instructions for *ACL authors. It contains instructions for using the \LaTeX{} style files for ACL conferences. 
-The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
-These instructions should be used both for papers submitted for review and for final versions of accepted papers.
+
 \end{abstract}
@@
 \section{Introduction}
-
-These instructions are for authors submitting papers to *ACL conferences using \LaTeX.
-They are not self-contained.
-All authors must follow the general instructions for *ACL proceedings,\footnote{\url{http://acl-org.github.io/ACLPUB/formatting.html}} and this document contains additional instructions for the \LaTeX{} style files.
-
-The templates include the \LaTeX{} source of this document (\texttt{acl.tex}),
-the \LaTeX{} style file used to format it (\texttt{acl.sty}),
-an ACL bibliography style (\texttt{acl\_natbib.bst}),
-an example bibliography (\texttt{custom.bib}),
-and the bibliography for the ACL Anthology (\texttt{anthology.bib}).
 
@@
 The metric is known to correlate directly with human sentence processing difficulty (Hale, 2001; Levy, 2008; Smith and Levy, 2013).
-In order to probe the model’s understanding of negative polarity licensing, we will construct a synthetic dataset which contained both grammatical examples, in which negative polarity items are correctly licensed, and ungrammatical examples, in which the licensor is either not present or is not in a c-command relationship with the licensee.
+In order to probe the model’s understanding of negative polarity licensing, we will construct a synthetic data set which contained both grammatical examples, in which negative polarity items are correctly licensed, and ungrammatical examples, in which the licensor is either not present or is not in a c-command relationship with the licensee.
 
-\section{Preamble}
+\section{Representation of Negative Polarity Items}
+Negative polarity items have a variety of properties.
+Most obviously, they should only be present in negative contexts.
+They are also immune to intervening material.
+One negative context can license multiple negative polarity items.
+Lastly, the negative item must be in the correct syntactic position to license a negative polarity item.
+In this section we demonstrate that LSTMs have learned these four properties of negative polarity items.
 
-The first line of the file must be
-\begin{quote}
-\begin{verbatim}
-\documentclass[11pt]{article}
-\end{verbatim}
-\end{quote}
+\subsection{Detection of Negative Polarity Items}
+As discussed in the introduction, negative polarity items are only grammatically correct in negative contexts.
+The following examples demonstrate this fact because only the sentences with negative contexts (a,c) are grammatically permissible:
+\begin{exe}
+[]
+\begin{xlist}
+\ex Jane didn't buy any books.
+\ex[*]{Jane did buy any books.}
+\ex He hasn't ever read Moby Dick.
+\ex[*]{He has ever read Moby Dick}
+\end{xlist}
+\end{exe}
+In this subsection we hope to demonstrate that LSTMs have expectations for negative polarity items.
+We designed 22 sentences with negaitve polarity items each in a positive and negative context.
+We expect to see higher surprisal metric in all of the positive context sentences.
+\subsection{Robustness of Negative Polarity Items to Intervening Material}
+Syntactic dependencies are unchanged by the presence intervening material.
+In the following example, the presence of a negative polarity item (anymore) is due to syntactic licensing by negation (don't); modifying the object doesn't change the sentence structure, so it has no effect on NPI licensing:
+\begin{exe}
+[]
+\begin{xlist}
+\ex We don't go to the supermarket anymore.
+\ex We don't go down the block to the supermarket that is owned by Mark anymore. 
+\end{xlist}
+\end{exe}
+We showed in the previous subsection that LSTMs have expectations for negative polarity items, so we attempt to determine whether those expectations remain over intervening material.
+For this subsection, we designed 10 sentences base sentences with NPI items.
+Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item.
+In all cases the intervening words modify the object of the base sentence.
+\subsection{Ability to License Multiple Negative Polarity Items}
+In most cases, a negative context can license multiple negative polarity items.
+\begin{exe}
+[]
+\begin{xlist}
+\ex He didn't want to go with anybody anywhere
+\ex [*]{He did want to go with anybody anywhere} 
+\end{xlist}
+\end{exe}
 
-To load the style file in the review version:
-\begin{quote}
-\begin{verbatim}
-\usepackage[review]{acl}
-\end{verbatim}
-\end{quote}
-For the final version, omit the \verb|review| option:
-\begin{quote}
-\begin{verbatim}
-\usepackage{acl}
-\end{verbatim}
-\end{quote}
-
-To use Times Roman, put the following in the preamble:
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\end{verbatim}
-\end{quote}
-(Alternatives like txfonts or newtx are also acceptable.)
-
-Please see the \LaTeX{} source of this document for comments on other packages that may be useful.
-
-Set the title and author using \verb|\title| and \verb|\author|.
-Within the author list, format multiple authors using \verb|\and| and \verb|\And| and \verb|\AND|; please see the \LaTeX{} source for examples.
-
-By default, the box containing the title and author names is set to the minimum of 5 cm. If you need more space, include the following in the preamble:
-\begin{quote}
-\begin{verbatim}
-\setlength\titlebox{<dim>}
-\end{verbatim}
-\end{quote}
-where \verb|<dim>| is replaced with a length. Do not set this length smaller than 5 cm.
-
+\subsection{Syntactic Position of Negative Context}
+Just because one section of a sentence is in a negative context does not mean that it is appropriate to use a negative polarity item in another clause.
+In the following example, the dependent clause is fixed as negative and the independent clause (which contains the NPI) is varied, so only the first variation is correct:
+\begin{exe}
+[]
+\begin{xlist}
+\ex Because he was not tired, he was not even sleeping
+\ex [*]{Because he was not tired, he was even sleeping} 
+\end{xlist}
+\end{exe}
+In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items.
+In order to do this we designed 10 base sentences.
+Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
 \section{Document Body}
 
-\subsection{Footnotes}
-
-Footnotes are inserted with the \verb|\footnote| command.\footnote{This is a footnote.}
-
-\subsection{Tables and figures}
-
-See Table~\ref{tab:accents} for an example of a table and its caption.
-\textbf{Do not override the default caption sizes.}
-
-\begin{table}
-\centering
-\begin{tabular}{lc}
-\hline
-\textbf{Command} & \textbf{Output}\\
-\hline
-\verb|{\"a}| & {\"a} \\
-\verb|{\^e}| & {\^e} \\
-\verb|{\`i}| & {\`i} \\ 
-\verb|{\.I}| & {\.I} \\ 
-\verb|{\o}| & {\o} \\
-\verb|{\'u}| & {\'u}  \\ 
-\verb|{\aa}| & {\aa}  \\\hline
-\end{tabular}
-\begin{tabular}{lc}
-\hline
-\textbf{Command} & \textbf{Output}\\
-\hline
-\verb|{\c c}| & {\c c} \\ 
-\verb|{\u g}| & {\u g} \\ 
-\verb|{\l}| & {\l} \\ 
-\verb|{\~n}| & {\~n} \\ 
-\verb|{\H o}| & {\H o} \\ 
-\verb|{\v r}| & {\v r} \\ 
-\verb|{\ss}| & {\ss} \\
-\hline
-\end{tabular}
-\caption{Example commands for accented characters, to be used in, \emph{e.g.}, Bib\TeX{} entries.}
-\label{tab:accents}
-\end{table}
-
-\subsection{Hyperlinks}
-
-Users of older versions of \LaTeX{} may encounter the following error during compilation: 
-\begin{quote}
-\tt\verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.
-\end{quote}
-This happens when pdf\LaTeX{} is used and a citation splits across a page boundary. The best way to fix this is to upgrade \LaTeX{} to 2018-12-01 or later.
-
-\subsection{Citations}
-
-\begin{table*}
-\centering
-\begin{tabular}{lll}
-\hline
-\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
-\hline
-\citep{Gusfield:97} & \verb|\citep| & \verb|\cite| \\
-\citealp{Gusfield:97} & \verb|\citealp| & no equivalent \\
-\citet{Gusfield:97} & \verb|\citet| & \verb|\newcite| \\
-\citeyearpar{Gusfield:97} & \verb|\citeyearpar| & \verb|\shortcite| \\
-\hline
-\end{tabular}
-\caption{\label{citation-guide}
-Citation commands supported by the style file.
-The style is based on the natbib package and supports all natbib citation commands.
-It also supports commands defined in previous ACL style files for compatibility.
-}
-\end{table*}
-
-Table~\ref{citation-guide} shows the syntax supported by the style files.
-We encourage you to use the natbib styles.
-You can use the command \verb|\citet| (cite in text) to get ``author (year)'' citations, like this citation to a paper by \citet{Gusfield:97}.
-You can use the command \verb|\citep| (cite in parentheses) to get ``(author, year)'' citations \citep{Gusfield:97}.
-You can use the command \verb|\citealp| (alternative cite without parentheses) to get ``author, year'' citations, which is useful for using citations within parentheses (e.g. \citealp{Gusfield:97}).
-
-\subsection{References}
-
-\nocite{Ando2005,borschinger-johnson-2011-particle,andrew2007scalable,rasooli-tetrault-2015,goodman-etal-2016-noise,harper-2014-learning}
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
-If your own bib file is named \texttt{custom.bib}, then placing the following before any appendices in your \LaTeX{} file will generate the references section for you:
-\begin{quote}
-\begin{verbatim}
-\bibliographystyle{acl_natbib}
-\bibliography{custom}
-\end{verbatim}
-\end{quote}
-
-You can obtain the complete ACL Anthology as a Bib\TeX{} file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
-To include both the Anthology and your own .bib file, use the following instead of the above.
-\begin{quote}
-\begin{verbatim}
-\bibliographystyle{acl_natbib}
-\bibliography{anthology,custom}
-\end{verbatim}
-\end{quote}
-
-Please see Section~\ref{sec:bibtex} for information on preparing Bib\TeX{} files.
-
-\subsection{Appendices}
-
-Use \verb|\appendix| before any appendix section to switch the section numbering over to letters. See Appendix~\ref{sec:appendix} for an example.
-
-\section{Bib\TeX{} Files}
-\label{sec:bibtex}
-
-Unicode cannot be used in Bib\TeX{} entries, and some ways of typing special characters can disrupt Bib\TeX's alphabetization.
-The recommended way of typing special characters is shown in Table~\ref{tab:accents}.
-
-Please ensure that Bib\TeX{} records contain DOIs or URLs when possible, and for all the ACL materials that you reference.
-Use the \verb|doi| field for DOIs and the \verb|url| field for URLs.
-If a Bib\TeX{} entry has a URL or DOI field, the paper title in the references section will appear as a hyperlink to the paper, using the hyperref \LaTeX{} package.
-
-\section*{Acknowledgements}
-
-This document has been adapted
-by Steven Bethard, Ryan Cotterell and Rui Yan
-from the instructions for earlier ACL and NAACL proceedings, including those for 
-ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
-NAACL 2019 by Stephanie Lukin and Alla Roskovskaya, 
-ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu, 
-NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
-Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
-ACL 2017 by Dan Gildea and Min-Yen Kan, 
-NAACL 2017 by Margaret Mitchell, 
-ACL 2012 by Maggie Li and Michael White, 
-ACL 2010 by Jing-Shin Chang and Philipp Koehn, 
-ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui, 
-ACL 2005 by Hwee Tou Ng and Kemal Oflazer, 
-ACL 2002 by Eugene Charniak and Dekang Lin, 
-and earlier ACL and EACL formats written by several people, including
-John Chen, Henry S. Thompson and Donald Walker.
-Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.
-
-% Entries for the entire Anthology, followed by custom entries
-\bibliography{anthology,custom}
-\bibliographystyle{acl_natbib}
-
-\appendix
-
-\section{Example Appendix}
-\label{sec:appendix}
-
-This is an appendix.
-
 \end{document}

### 2021-12-17 src/data.py
Update dictionary with <bos>

@@ -30,7 +30,7 @@ class Corpus(object):
         # Add words to the dictionary
         with open(path, 'r', encoding="utf8") as f:
             for line in f:
-                words = line.split() + ['<eos>']
+                words = ['<bos>'] + line.split() + ['<eos>']
                 for word in words:
                     self.dictionary.add_word(word)
 
@@ -38,7 +38,7 @@ class Corpus(object):
         with open(path, 'r', encoding="utf8") as f:
             idss = []
             for line in f:
-                words = line.split() + ['<eos>']
+                words = ['<bos>'] + line.split() + ['<eos>']
                 ids = []
                 for word in words:
                     ids.append(self.dictionary.word2idx[word])

### 2021-12-17 src/train.py
Retrain model with batch_size 128

@@ -27,6 +27,9 @@ parser.add_argument('--epochs', type=int, default=40,
                     help='upper epoch limit')
 parser.add_argument('--batch_size', type=int, default=128, metavar='N',
                     help='batch size')
+parser.add_argument('--micro_batch_size', type=int, default=32, metavar='N',
+                    help='To save GPU memory, we split up batches into micro\
+                    batches.')
 parser.add_argument('--bptt', type=int, default=35,
                     help='sequence length')
 parser.add_argument('--dropout', type=float, default=0.2,
@@ -46,6 +49,8 @@ parser.add_argument('--dry-run', action='store_true',
 
 args = parser.parse_args()
 
+assert args.batch_size % args.micro_batch_size == 0
+micro_batches_per_batch = args.batch_size // args.micro_batch_size
 
 # Set the random seed manually for reproducibility.
 torch.manual_seed(args.seed)
@@ -84,7 +89,7 @@ def batchify(data, bsz):
     return data.to(device)
 
 eval_batch_size = 10
-train_data = batchify(corpus.train, args.batch_size)
+train_data = batchify(corpus.train, args.micro_batch_size)
 val_data = batchify(corpus.valid, eval_batch_size)
 test_data = batchify(corpus.test, eval_batch_size)
 
@@ -139,7 +144,7 @@ def evaluate(data_source):
             data, targets = get_batch(data_source, i)
             output, hidden = model(data, hidden)
             hidden = repackage_hidden(hidden)
-            total_loss += len(data) * criterion(output, targets).item()
+            total_loss += len(data) * criterion(output.view(-1, ntokens), targets.view(-1)).item()
     return total_loss / (len(data_source) - 1)
 
 
@@ -149,32 +154,36 @@ def train():
     total_loss = 0.
     start_time = time.time()
     ntokens = len(corpus.dictionary)
-    hidden = model.init_hidden(args.batch_size)
-    for batch, i in enumerate(range(0, train_data.size(0) - 1, args.bptt)):
+    hidden = model.init_hidden(args.micro_batch_size)
+    model.zero_grad()
+    for micro_batch, i in enumerate(range(0, train_data.size(0) - 1, args.bptt)):
+        batch = micro_batch // micro_batches_per_batch
         data, targets = get_batch(train_data, i)
         # Starting each batch, we detach the hidden state from how it was
         # previously produced. If we didn't, the model would try backpropagating
         # all the way to start of the dataset.
[…]

### 2021-12-17 src/main.py
Fix bugs

@@ -13,7 +13,7 @@ import model
 parser = argparse.ArgumentParser(description='Evaluate language model on examples')
 parser.add_argument('--model', type=str, default='lstm.pt',
                     help="location of model to evaluate")
-parser.add_argument('--evaluate', type=str, 
+parser.add_argument('--evaluate', type=str, default='../sentences.txt',
                     help='location of example sentences to evaluate')
 parser.add_argument('--dictionary', type=str, default='dictionary.p',
                     help='location of the dictionary to use')
@@ -41,6 +41,10 @@ with open(args.model, 'rb') as f:
     model.rnn.flatten_parameters()
     model.eval()
 
+def word_to_id(dictionary, word):
+    unk_id = dictionary.word2idx['<unk>']
+    return dictionary.word2idx.get(word, unk_id)
+    
 def read_examples_file_as_tensor(examples, dictionary):
     """Reads a text file with one example sentence on each line, and returns a
     tensor of shape (max_len + 1, num_sentences), where max_len is the length of
@@ -51,7 +55,7 @@ def read_examples_file_as_tensor(examples, dictionary):
         seq_len = 0
         for line in f:
             words = line.split() + ['<eos>']
-            ids = [dictionary.word2idx[w] for w in words]
+            ids = [word_to_id(dictionary, w) for w in words]
             sequences.append(ids)
             if len(words) > seq_len:
                 seq_len = len(words)
@@ -103,6 +107,7 @@ def plot_surprisal(vocab, model, batch):
         
         series = pd.Series(seq_surprisal, seq_words)
         series.plot()
+        plt.xticks(range(len(seq_words)), labels=seq_words)
         plt.show()
 
 

### 2021-12-17 src/main.py
Generate plots using seaborn

@@ -3,6 +3,7 @@ import os
 import pickle
 import pandas as pd
 import matplotlib.pyplot as plt
+import seaborn as sns
 
 import torch
 import torch.nn as nn
@@ -112,5 +113,6 @@ def plot_surprisal(vocab, model, batch):
 
 
 batch = read_examples_file_as_tensor(args.evaluate, dictionary)
+sns.set()
 plot_surprisal(dictionary, model, batch)
 

### 2021-12-18 paper.tex
Updates from Overleaf
section 3.1

@@
 \usepackage{gb4e}
+\noautomath
 % Standard package includes
@@
 \usepackage{latexsym}
+
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{subfig}
+\graphicspath{{./img/}}
 
@@
 
-% If the title and author information does not fit in the area allocated, uncomment the following
-%
-%\setlength\titlebox{<dim>}
-%
-% and set <dim> to something 5cm or larger.
-
-\title{Final Paper for LING 380}
-
-% Author information can be set in various styles:
-% For several authors from the same institution:
-% \author{Author 1 \and ... \and Author n \\
-%         Address line \\ ... \\ Address line}
-% if the names do not fit well on one line use
-%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
-% For authors from different institutions:
-% \author{Author 1 \\ Address line \\  ... \\ Address line
-%         \And  ... \And
-%         Author n \\ Address line \\ ... \\ Address line}
-% To start a seperate ``row'' of authors use \AND, as in
-% \author{Author 1 \\ Address line \\  ... \\ Address line
-%         \AND
-%         Author 2 \\ Address line \\ ... \\ Address line \And
-%         Author 3 \\ Address line \\ ... \\ Address line}
+\title{Do LSTMs Understand the Licensing of Negative Polarity Items?}
 
@@
 The metric is defined to be the log inverse probability:
-\[S(x_i) = -log_2(p_i|h_{i-1}),\]
-where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is the softmax activation. The surprisal is measured in bits because we take a base two logarithm.
+\[S(x_i) = -\log_2 P(x_i\,|\,h_{i-1}),\]
+where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is given by the softmax activation.
+The surprisal is measured in bits because we take a base two logarithm.
 
@@
 Lastly, the negative item must be in the correct syntactic position to license a negative polarity item.
-In this section we demonstrate that LSTMs have learned these four properties of negative polarity items.
+In this section we demonstrate that LSTMs have learned these four properties negative polarity items.
 
@@
 \begin{exe}
-[]
+\ex
 \begin{xlist}
-\ex Jane didn't buy any books.
-\ex[*]{Jane did buy any books.}
-\ex He hasn't ever read Moby Dick.
-\ex[*]{He has ever read Moby Dick}
+\ex Bill is not here yet.
+\ex[*]{Bill is here yet.}
+\ex They hired somebody without any talent.
+\ex[*]{They hired somebody with any talent.}
 \end{xlist}
@@
 In this subsection we hope to demonstrate that LSTMs have expectations for negative polarity items.
-We designed 22 sentences with negaitve polarity items each in a positive and negative context.
+We designed 20 sentences with negaitve polarity items each in a positive and negative context.
 We expect to see higher surprisal metric in all of the positive context sentences.
+
+The plots of surprisal in Figure 1 are two strong examples of the general trend that we uncovered; negative polarity items are much more surprising to the model when contained positive contexts rather than negative contexts.
+The negative polarity item in 19 of the 20 base sentences that we created caused higher surprisal when they were placed in a positive context.
+
+Overall, this experiment shows that our LSTM model has at least a basic understanding of when negative polarity items should be expected.
+
+\begin{figure}
+    \centering
+    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{good1} }}
+    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad1} }}
+    \qquad
+    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{good2} }}
+    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad2} }}
+    \caption{}
+    \label{fig:example}
+\end{figure}
+
 \subsection{Robustness of Negative Polarity Items to Intervening Material}
@@
 \begin{exe}
-[]
+\ex
 \begin{xlist}
@@
 
-\subsection{Syntactic Position of Negative Context}
+\subsection{Licensing of Negative Polarity Items Between Clauses}
 Just because one section of a sentence is in a negative context does not mean that it is appropriate to use a negative polarity item in another clause.

### 2021-12-18 paper.tex
Update section 3.2

@@
 The plots of surprisal in Figure 1 are two strong examples of the general trend that we uncovered; negative polarity items are much more surprising to the model when contained positive contexts rather than negative contexts.
-The negative polarity item in 19 of the 20 base sentences that we created caused higher surprisal when they were placed in a positive context.
+The negative polarity item in 19 of the 20 base sentences that we created were given higher surprisal scores when they were placed in a positive context.
 
@@
     \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad2} }}
-    \caption{}
+    \caption{Plots of surprisal scores for words in the example sentences.
+Note that plots (a) and (c) are both contextually negative and plots (b) and (d) are positive.
+As expected, the surprisal score for the NPIs (\textit{yet} in plots (a) and (b) and \textit{any} in plots (c) and (d)) is much higher in positive contexts.}
     \label{fig:example}
@@
 \begin{xlist}
-\ex We don't go to the supermarket anymore.
-\ex We don't go down the block to the supermarket that is owned by Mark anymore. 
+\ex We do not go to the supermarket anymore.
+\ex We do not go to the supermarket that is owned by Mark anymore.
+\ex We do not go down the block to the supermarket that is owned by Mark anymore. 
 \end{xlist}
@@
 In all cases the intervening words modify the object of the base sentence.
-\subsection{Ability to License Multiple Negative Polarity Items}
-In most cases, a negative context can license multiple negative polarity items.
-\begin{exe}
-\ex
-\begin{xlist}
-\ex He didn't want to go with anybody anywhere
-\ex [*]{He did want to go with anybody anywhere} 
-\end{xlist}
-\end{exe}
 
-\subsection{Licensing of Negative Polarity Items Between Clauses}
+\begin{figure*}
+    \centering
+    \subfloat[Base Sentence]{{\includegraphics[width=.32\textwidth]{short} }}
+    \subfloat[5 Additional Words]{{\includegraphics[width=.32\textwidth]{medium} }}
+    \subfloat[8 Additional Words]{{\includegraphics[width=.32\textwidth]{long} }}
+    \caption{Plots of surprisal scores for words in the modified sentence.
+The surprisal score of the negative polarity item \textit{anymore} remains relatively unchanged by the amount of intervening material.}
+\end{figure*}
+
+An example of the results for this section are visualized in Figure 2.
+We found that in general, there was a weak positive correlation between the number of words in between the word that determined the negative context and the negative polarity item and the surprisal score.
+However as seen in the plots, the scores are relatively similar, so we can conclude that the model does not consider the NPIs as ungrammatical when the amount of intervening words is long.
+\subsection{Licensing of Negative Polarity Items Between Separate C-Commands}
 Just because one section of a sentence is in a negative context does not mean that it is appropriate to use a negative polarity item in another clause.
@@
 Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
+We are motivated to analyze this type of interaction because a model that is well adjusted to NPI should determine that the presence of negative context in a separate c-command does not influence whether it is actually grammatical.
+\subsection{Ability to License Multiple Negative Polarity Items}
+In most cases, a negative context can license multiple negative polarity items.
+\begin{exe}
+\ex
+\begin{xlist}
+\ex He didn't want to go with anybody anywhere
+\ex [*]{He did want to go with anybody anywhere} 
+\end{xlist}
+\end{exe}
+
 \section{Document Body}

### 2021-12-21 src/main.py
Define surprisal metric function

@@ -53,10 +53,18 @@ def read_examples_file_as_tensor(examples, dictionary):
     number of sentences."""
     with open(examples, 'r', encoding="utf8") as f:
         sequences = []
+        npi_marker_indexes = []
         seq_len = 0
         for line in f:
             words = line.split() + ['<eos>']
-            ids = [word_to_id(dictionary, w) for w in words]
+
+            if '*' not in words:
+                print("Skipping sentence with no NPI marker")
+                continue
+            
+            npi_marker_indexes.append(words.index('*'))
+            
+            ids = [word_to_id(dictionary, w) for w in words if w != '*']
             sequences.append(ids)
             if len(words) > seq_len:
                 seq_len = len(words)
@@ -64,7 +72,7 @@ def read_examples_file_as_tensor(examples, dictionary):
     eos_id = dictionary.word2idx['<eos>']
     sequences = [s + [eos_id] * (seq_len - len(s)) for s in sequences]
     sequences = [torch.tensor(seq).type(torch.int64) for seq in sequences]
-    return torch.stack(sequences)
+    return torch.stack(sequences), npi_marker_indexes
     
 def surprisal(model, batch):
     """Calculates the surprisal for each word (with the first word having
@@ -117,7 +125,21 @@ def plot_surprisal(vocab, model, batch):
         plt.show()
 
 
-batch = read_examples_file_as_tensor(args.evaluate, dictionary)
+def compute_surprisal_on_batch(model, batch, markers):
+    surp = surprisal(model, batch)
+    point_surprisals = []
+    cumulative_surprisals = []
+    for i, sentence in enumerate(batch):
+        marker = markers[i]
+        point_surprisal = surp[i, marker]
+        cumulative_surprisal = 0
+        for j in range(marker, sentence.size(0)):
+            cumulative_surprisal += surp[i, j]
+        point_surprisals.append(point_surprisal)
+        cumulative_surprisals.append(cumulative_surprisal)
+    return point_surprisals, cumulative_surprisals
+
+batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
 sns.set()
 plot_surprisal(dictionary, model, batch)
 

### 2021-12-21 src/main.py
Remove EOS tokens from plots

@@ -104,6 +104,11 @@ def plot_surprisal(vocab, model, batch):
     for i, seq in enumerate(batch):
         seq_surprisal = surp[i]
         seq_surprisal = seq_surprisal.detach().numpy()
+
+        seq_end = seq.index(vocab.word2idx['<eos>'])
+        seq = seq[0:seq_end]
+        seq_surprisal = seq_surprisal[0:seq_end]
+
         seq_words = [vocab.idx2word[word_id] for word_id in seq.tolist()]
         
         series = pd.Series(seq_surprisal, seq_words)

### 2021-12-21 src/main.py
Changes

@@ -56,7 +56,13 @@ def read_examples_file_as_tensor(examples, dictionary):
         npi_marker_indexes = []
         seq_len = 0
         for line in f:
-            words = line.split() + ['<eos>']
+            if not line:
+                continue
+
+            words = ['<bos>'] + line.split() + ['<eos>']
+
+            if words[1] == '#':
+                continue
 
             if '*' not in words:
                 print("Skipping sentence with no NPI marker")
@@ -113,7 +119,7 @@ def plot_surprisal(vocab, model, batch):
         seq_surprisal = surp[i]
         seq_surprisal = seq_surprisal.detach().numpy()
 
-        seq_end = seq.index(vocab.word2idx['<eos>'])
+        seq_end = seq.tolist().index(vocab.word2idx['<eos>']) + 1
         seq = seq[0:seq_end]
         seq_surprisal = seq_surprisal[0:seq_end]
 
@@ -122,22 +128,62 @@ def plot_surprisal(vocab, model, batch):
         series = pd.Series(seq_surprisal, seq_words)
         series.plot()
         plt.xticks(range(len(seq_words)), labels=seq_words)
+
         plt.show()
 
 
-def compute_surprisal_on_batch(model, batch, markers):
+def compute_surprisal_on_batch(model, dictionary, batch, markers):
     surp = surprisal(model, batch)
-    point_surprisals = []
-    cumulative_surprisals = []
-    for i, sentence in enumerate(batch):
+    results = []
+    for i, seq in enumerate(batch):
+        seq_surprisal = surp[i]
         marker = markers[i]
-        point_surprisal = surp[i, marker]
+
+        seq_end = seq.tolist().index(dictionary.word2idx['<eos>']) + 1
+        seq = seq[0:seq_end]
+        seq_surprisal = seq_surprisal[0:seq_end]
+        
+        
+        point_surprisal = surp[i, marker].item()
+
         cumulative_surprisal = 0
-        for j in range(marker, sentence.size(0)):
-            cumulative_surprisal += surp[i, j]
-        point_surprisals.append(point_surprisal)
-        cumulative_surprisals.append(cumulative_surprisal)
-    return point_surprisals, cumulative_surprisals
+        for j in range(marker, seq.size(0)):
+            cumulative_surprisal += surp[i, j].item()
+
+        results.append((point_surprisal, cumulative_surprisal))
+        
+    return results
+
[…]

### 2021-12-21 src/main.py
Continue_seq function

@@ -183,7 +183,21 @@ def generate_seq(model, dictionary, length):
         result.append(dictionary.idx2word[seq[0, i]])
 
     return result
-        
+
+def continue_seq(model, dictionary, seq, choices):
+    seq = [dictionary.word2idx[w] for w in seq]
+    seq = torch.tensor([seq])
+
+    seq_len = seq.size(-1)
+    hidden = model.init_hidden(seq_len)
+
+    logits = model(seq, hidden)[0]
+    idx = torch.argsort(logits, descending=True)
+    out = idx[:, seq_len - 1, :choices]
+    out = out.squeeze().tolist()
+    out = [dictionary.idx2word[i] for i in out]
+
+    return out
 
 batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
 sns.set()

### 2021-12-21 paper.tex
Updates from Overleaf

@@
 
-\documentclass[11pt]{article}
+\documentclass[11pt, round]{article}
 
@@
 \usepackage{caption}
-\usepackage{subcaption}
 \usepackage{subfig}
@@
 
+\usepackage{natbib}
+\bibliographystyle{plainnat}
+
 \title{Do LSTMs Understand the Licensing of Negative Polarity Items?}
@@
 \maketitle
+
 \begin{abstract}
-
+While LSTMs have proven themselves as powerful tools for language modeling tasks, the extent to which their generalizations are based on underlying syntactic structure remains an open area of inquiry.
+We focus our investigation on the relationship between negative polarity items and their licensors in English.
+We probe the model's representation of this relationship by computing the network's surprisal upon encountering NPIs in licensed and unlicensed contexts.
+We show that...
 \end{abstract}
@@
 
+The development of Recurrent Neural Networks (RNN) has been a recent hallmark of advancements in Natural Language Processing (NLP).
+The inherent ability for RNNs to learn and internalize a number of linguistic capabilities is evidenced by a large corpus of research in the utilization of RNNs for NLP tasks such as machine translation, language modeling, and syntactic parsing \cite{wilcox-etal-2018-rnn}.
+However, it is still largely unknown how such networks are able to deduce linguistic information, in that there is comparatively less literature present regarding if networks' success is due to opportunistic surface-pattern-based heuristics or if networks are able to genuinely extract and interpret structural properties, syntactic or semantic, of language from their training data \cite{lakretz2019emergence}.
+
+One subset of RNNs, namely Long Short-Term Memory RNNs (LSTMs), has grown in utility in said extraction of structural properties of language.
+Wilcox et al.
+\shortcite{wilcox-etal-2018-rnn} determined utilized LSTM models to explore the syntactic relationship of the filler-gap dependency, the relationship between a filler word (in English, a wh- complementizer like "what" or "when") and a gap word (in English, an empty syntactic position uniquely licensed by a filler word).
+Lakretz et al.
+\shortcite{lakretz2019emergence} similarly utilized LSTM models for exploring numerical agreement between verbs and their arguments.
+
+There is, though, little research investigating the feature of licensing of negative polarity items in language.
+Negative polarity items are a class of words that are only permitted to appear in certain “negative” contexts.
+For example, in English, negative polarity items include words like "ever," "any," and "even."
+\begin{enumerate}
+    \item He never gave up.
+    \item *He \textit{ever} gave up.
+    \item Nobody ever gave up.
+\end{enumerate}
+Here, the word "ever" is not permitted to occur in a sentence except when licensed by a word such as "nobody" that provides the surrounding "negative" context.
+Notably, the relationship between the licensor and the negative polarity item is based on the hierarchical structure of a sentence, wherein the licensor must c-command the negative polarity item (i.e.
+the negative polarity item must be a sister of the licensor or a direct descendant of a sister node of the licensor).
+
+If a language model, after being trained on a sentence data set implementing the usage of negative polarity licensing, is able to predict the distribution of negative polarity items and the contexts in which they are licensed with significant accuracy, this would be evidence that the model is able to learn this feature of syntax, in particular that it is able to detect a relationship between a licensor and a negative polarity item.
+
 \section{Methods}
 \subsection{Language Models}
+
+For our investigation, we trained a LSTM model on the WikiText-2 dataset \cite{merity2016pointer}, which consists of over 2 million tokens from 600 Wikipedia articles, with a vocabulary size of 33,279.
+Our LSTM has an embedding layer, assigning 200-dimensional embeddings to each word in the vocabulary, and two recurrent layers, each with 650 hidden units.
+During training, we use dropout regularization as described in Srivastava et.
+al.
+\shortcite{srivastava2014dropout}, with a dropout rate set at 0.2.
+Optimization was performed using gradient descent, with a base learning rate of 0.2, which we annealed whenever performance did not improve on the development set.
+
 \subsection{Surprisal}
-We asses the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric that is defined in Wilcox et.
+
+We assess the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric that, as used in Wilcox et.
 al.
-(2018).
+\shortcite{wilcox-etal-2018-rnn}.
 The metric is defined to be the log inverse probability:
-\[S(x_i) = -\log_2 P(x_i\,|\,h_{i-1}),\]
+\[ S(x_i) = -\log_2 P(x_i \, | \, h_{i-1}), \]
 where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is given by the softmax activation.
@@
 
-A high surprisal metric means that a given word was unexpected in its context.
-The metric is known to correlate directly with human sentence processing difficulty (Hale, 2001; Levy, 2008; Smith and Levy, 2013).
-In order to probe the model’s understanding of negative polarity licensing, we will construct a synthetic data set which contained both grammatical examples, in which negative polarity items are correctly licensed, and ungrammatical examples, in which the licensor is either not present or is not in a c-command relationship with the licensee.
+A high surprisal metric means that, in its context, a word was predicted with \textit{low} probability, and therefore was unexpected by the model.
+A direct correlation has been shown between language model surprisal metrics and  human sentence processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+In order to probe the model’s understanding of negative polarity licensing, we construct a synthetic data set which contained both grammatical examples, in which negative polarity items are correctly licensed, and ungrammatical examples, in which the licensor is either not present or is not in a c-command relationship with the licensee.
+
+\subsection{Experimental Design}
 
 \section{Representation of Negative Polarity Items}
+
 Negative polarity items have a variety of properties.
@@
 
-\subsection{Detection of Negative Polarity Items}
+\subsection{Licensing of Negative Polarity Items}
+
 As discussed in the introduction, negative polarity items are only grammatically correct in negative contexts.
@@
 \begin{exe}
-\ex
+\ex\label{ex:license-1}
 \begin{xlist}
@@
 \ex[*]{Bill is here yet.}
+\end{xlist}
+\ex\label{ex:license-2}
+\begin{xlist}
 \ex They hired somebody without any talent.
@@
 \end{exe}
-In this subsection we hope to demonstrate that LSTMs have expectations for negative polarity items.
-We designed 20 sentences with negaitve polarity items each in a positive and negative context.
-We expect to see higher surprisal metric in all of the positive context sentences.
+In this subsection, we confirm the hypothesis that surprisal metrics for NPIs are reduced in the presence of an appropriate licensor.
+We created 20 test items containing negative polarity items.
+For each item, we constructed two variants: one with a licensor, and one without a licensor, as exemplified in Examples \ref{ex:license-1} and \ref{ex:license-2}.
 
-The plots of surprisal in Figure 1 are two strong examples of the general trend that we uncovered; negative polarity items are much more surprising to the model when contained positive contexts rather than negative contexts.
-The negative polarity item in 19 of the 20 base sentences that we created were given higher surprisal scores when they were placed in a positive context.
+Plotted in Figure \ref{fig:surprisal-basic-licensing} are graphs of surprisal on two generic examples; negative polarity items are assigned lower surprisal values by the model when they are found in grammatically licit environments.
 
@@
 As expected, the surprisal score for the NPIs (\textit{yet} in plots (a) and (b) and \textit{any} in plots (c) and (d)) is much higher in positive contexts.}
-    \label{fig:example}
+    \label{fig:surprisal-basic-licensing}
 \end{figure}
@@
 
-\section{Document Body}
+\section{Results}
+
+\bibliography{sources}
 

### 2021-12-21 paper.tex
Updates from Overleaf

@@
 The inherent ability for RNNs to learn and internalize a number of linguistic capabilities is evidenced by a large corpus of research in the utilization of RNNs for NLP tasks such as machine translation, language modeling, and syntactic parsing \cite{wilcox-etal-2018-rnn}.
+
 However, it is still largely unknown how such networks are able to deduce linguistic information, in that there is comparatively less literature present regarding if networks' success is due to opportunistic surface-pattern-based heuristics or if networks are able to genuinely extract and interpret structural properties, syntactic or semantic, of language from their training data \cite{lakretz2019emergence}.
@@
 Wilcox et al.
-\shortcite{wilcox-etal-2018-rnn} determined utilized LSTM models to explore the syntactic relationship of the filler-gap dependency, the relationship between a filler word (in English, a wh- complementizer like "what" or "when") and a gap word (in English, an empty syntactic position uniquely licensed by a filler word).
+\shortcite{wilcox-etal-2018-rnn} determined utilized LSTM models to explore the syntactic relationship of the filler-gap dependency, the relationship between a filler word (in English, a wh- complementizer like \textit{what} or \textit{when}) and a gap word (in English, an empty syntactic position uniquely licensed by a filler word).
 Lakretz et al.
@@
 There is, though, little research investigating the feature of licensing of negative polarity items in language.
-Negative polarity items are a class of words that are only permitted to appear in certain “negative” contexts.
-For example, in English, negative polarity items include words like "ever," "any," and "even."
+Negative polarity items are a class of words that are only permitted to appear in certain "negative" contexts.
+For example, in English, negative polarity items include words like \textit{ever}, \textit{any}, and \textit{even}.
 \begin{enumerate}
@@
 \end{enumerate}
-Here, the word "ever" is not permitted to occur in a sentence except when licensed by a word such as "nobody" that provides the surrounding "negative" context.
+Here, the word \textit{ever} is not permitted to occur in a sentence except when licensed by a word such as \textit{nobody} that provides the surrounding "negative" context.
 Notably, the relationship between the licensor and the negative polarity item is based on the hierarchical structure of a sentence, wherein the licensor must c-command the negative polarity item (i.e.

### 2021-12-21 src/main.py
Changes

@@ -167,22 +167,13 @@ def compute_licensing_interaction(surprisal_list):
 
     return results
 
-def generate_seq(model, dictionary, length):
-    seq = ['The', 'moon', 'is', 'a']
-    seq = [dictionary.word2idx[w] for w in seq]
-    seq = torch.tensor([seq])
-    
+def generate_seq(model, dictionary, seed, length):
+    seq = seed.split()
     for i in range(length):
-        hidden = model.init_hidden(i + 5)
-        logits = model(seq, hidden)[0][:, i]
-        out = logits.argmax(dim=-1, keepdims=True)
-        seq = torch.cat([seq, out], dim=-1)
+        next_word = continue_seq(model, dictionary, seq, 2)[0]
+        seq.append(next_word)
         
-    result = []
-    for i in range(seq.size(-1)):
-        result.append(dictionary.idx2word[seq[0, i]])
-
-    return result
+    return seq
 
 def continue_seq(model, dictionary, seq, choices):
     seq = [dictionary.word2idx[w] for w in seq]
@@ -199,7 +190,22 @@ def continue_seq(model, dictionary, seq, choices):
 
     return out
 
+def foo():
+    batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
+    thing = compute_surprisal_on_batch(model, dictionary, batch, npi_markers)
+    thing = compute_licensing_interaction(thing)
+    thing = [j for i, j in thing]
+    return ttest_1samp(thing, 0, alternative='greater')
+
+def bar():
+    batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
+    thing = compute_surprisal_on_batch(model, dictionary, batch, npi_markers)
+    thing = compute_licensing_interaction(thing)
+    thing = [j for i, j in thing]
+    thing = pd.Series(thing)
+    return thing
+    
+    
 batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
 sns.set()
 plot_surprisal(dictionary, model, batch)
-

### 2021-12-21 src/train.py

@@ -129,7 +129,7 @@ def repackage_hidden(h):
 def get_batch(source, i):
     seq_len = min(args.bptt, len(source) - 1 - i)
     data = source[i:i+seq_len]
-    target = source[i+1:i+1+seq_len]
+    target = source[i+1:i+1+seq_len].view(-1)
     return data, target
 
 
@@ -143,8 +143,9 @@ def evaluate(data_source):
         for i in range(0, data_source.size(0) - 1, args.bptt):
             data, targets = get_batch(data_source, i)
             output, hidden = model(data, hidden)
+            output = output.view(-1, ntokens)
+            total_loss += criterion(output, targets, reduction='sum').item()
             hidden = repackage_hidden(hidden)
-            total_loss += len(data) * criterion(output.view(-1, ntokens), targets.view(-1)).item()
     return total_loss / (len(data_source) - 1)
 
 
@@ -164,7 +165,8 @@ def train():
         # all the way to start of the dataset.
         hidden = repackage_hidden(hidden)
         output, hidden = model(data, hidden)
-        loss = criterion(output.view(-1, ntokens), targets.view(-1))
+        output = output.view(-1, ntokens)
+        loss = criterion(output, targets)
         loss.backward()
 
         # `clip_grad_norm` helps prevent the exploding gradient problem in RNNs

### 2021-12-21 src/main.py
Fixes

@@ -154,7 +154,7 @@ def compute_surprisal_on_batch(model, dictionary, batch, markers):
         
     return results
 
-def compute_licensing_interaction(surprisal_list):
+def compute_licensing_interaction_paired(surprisal_list):
     good = surprisal_list[::2]
     bad = surprisal_list[1::2]
     
@@ -193,14 +193,14 @@ def continue_seq(model, dictionary, seq, choices):
 def foo():
     batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
     thing = compute_surprisal_on_batch(model, dictionary, batch, npi_markers)
-    thing = compute_licensing_interaction(thing)
+    thing = compute_licensing_interaction_paired(thing)
     thing = [j for i, j in thing]
     return ttest_1samp(thing, 0, alternative='greater')
 
 def bar():
     batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
     thing = compute_surprisal_on_batch(model, dictionary, batch, npi_markers)
-    thing = compute_licensing_interaction(thing)
+    thing = compute_licensing_interaction_paired(thing)
     thing = [j for i, j in thing]
     thing = pd.Series(thing)
     return thing

### 2021-12-21 paper.tex

@@
 The development of Recurrent Neural Networks (RNN) has been a recent hallmark of advancements in Natural Language Processing (NLP).
-The inherent ability for RNNs to learn and internalize a number of linguistic capabilities is evidenced by a large corpus of research in the utilization of RNNs for NLP tasks such as machine translation, language modeling, and syntactic parsing \cite{wilcox-etal-2018-rnn}.
+The inherent ability for RNNs to learn and internalize a number of linguistic capabilities is evidenced by a large corpus of research in the utilization of RNNs for NLP tasks such as machine translation, language modeling, and syntactic parsing \cite{wilcox-etal-2018-rnn,gulordava2018colorless,jozefowicz2016exploring}.
 
-However, it is still largely unknown how such networks are able to deduce linguistic information, in that there is comparatively less literature present regarding if networks' success is due to opportunistic surface-pattern-based heuristics or if networks are able to genuinely extract and interpret structural properties, syntactic or semantic, of language from their training data \cite{lakretz2019emergence}.
+However, it is still largely unknown how such networks represent linguistic information internally.
+Historically, the majority of research on these networks has focused simply on evaluating and optimizing performance on applied NLP tasks.
+However, there has been a recent increase in literature probing inside of these black box models and evaluating the extent to which their success is based on surface level heuristics rather than genuinely extracting and interpreting structural properties of language, either syntactic or semantic.
+\cite{lakretz2019emergence}.
 
@@
 Wilcox et al.
-\shortcite{wilcox-etal-2018-rnn} determined utilized LSTM models to explore the syntactic relationship of the filler-gap dependency, the relationship between a filler word (in English, a wh- complementizer like \textit{what} or \textit{when}) and a gap word (in English, an empty syntactic position uniquely licensed by a filler word).
+\shortcite{wilcox-etal-2018-rnn} researched the ability of LSTM models to learn the syntactic filler-gap dependency, the relationship between a filler word (in English, a wh- complementizer like \textit{what} or \textit{when}) and a gap word (in English, an empty syntactic position uniquely licensed by a filler word).
 Lakretz et al.
-\shortcite{lakretz2019emergence} similarly utilized LSTM models for exploring numerical agreement between verbs and their arguments.
+\shortcite{lakretz2019emergence} similarly explored the ability of LSTM models to identify numerical agreement between verbs and their arguments.
 
@@
 
-If a language model, after being trained on a sentence data set implementing the usage of negative polarity licensing, is able to predict the distribution of negative polarity items and the contexts in which they are licensed with significant accuracy, this would be evidence that the model is able to learn this feature of syntax, in particular that it is able to detect a relationship between a licensor and a negative polarity item.
+If a language model after being trained on a generic linguistic data set is able to predict the distribution of negative polarity items and the contexts in which they are licensed with significant accuracy, this would be evidence that the model is able to learn this feature of syntax, in particular that it is able to detect a relationship between a licensor and a negative polarity item.
 
 \section{Methods}
-\subsection{Language Models}
+
+\subsection{Language Model}
 
@@
 \shortcite{srivastava2014dropout}, with a dropout rate set at 0.2.
-Optimization was performed using gradient descent, with a base learning rate of 0.2, which we annealed whenever performance did not improve on the development set.
+Optimization was performed using gradient descent, with a base learning rate of 0.2, which decreased exponentially whenever performance on the development set regressed.
 
@@
 
-We assess the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric that, as used in Wilcox et.
+We assess the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric, as applied in Wilcox et.
 al.
@@
 
-A high surprisal metric means that, in its context, a word was predicted with \textit{low} probability, and therefore was unexpected by the model.
+A high surprisal metric means that, in the given context, a word was predicted with \textit{low} probability, and was therefore unexpected by the model.
+Basically, given a hidden state of the model at a current time step, the surprisal metric is a way to measure the whether the next word makes \textit{sense}.
 A direct correlation has been shown between language model surprisal metrics and  human sentence processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-In order to probe the model’s understanding of negative polarity licensing, we construct a synthetic data set which contained both grammatical examples, in which negative polarity items are correctly licensed, and ungrammatical examples, in which the licensor is either not present or is not in a c-command relationship with the licensee.
 
@@
 
-Negative polarity items have a variety of properties.
-Most obviously, they should only be present in negative contexts.
-They are also immune to intervening material.
-One negative context can license multiple negative polarity items.
-Lastly, the negative item must be in the correct syntactic position to license a negative polarity item.
+Negative polarity items have a variety of linguistic properties which we would like to probe.
+On the most basic level, they are only allowed to occur in negative contexts.
+More specifically, in order for an NPI to be licit, it must be c-commanded by its licensor.
+Another property is that NPIs are immune to intervening material separating the NPI from its licensor.
+And finally, one negative context can license multiple negative polarity items.
 In this section we demonstrate that LSTMs have learned these four properties negative polarity items.
+
+Our general approach to studying these properties of the dependency between NPI and licensor is as follows.
+We construct a number of test items containing negative polarity items, including variation in whichever independent variable we are curious about (e.g.
+whether or not the licensor c-commands the NPI).
+For each item, we construct two variants: one with a licensor, and one without a licensor.
+We define the licensing interaction to be the difference in surprisals between these two variants:
+\[ \textrm{interaction} = S_{\textrm{no license}} - S_{\textrm{license}}. \]
+A large licensing interaction suggests that, in a given test item, the presence of a licensor does dampen the model's surprisal at encountering an NPI, while a small licensing interactoin suggests that, in a given test item, the licensor does not dampen the surprisal.
+
+In computing the licensing interaction, we measure surprisal in two different ways.
+First of all, we can measure the instantaneous surprisal of the model upon encountering an NPI.
+Secondly, we measure cumulative surprisal from the NPI to the end of its clause.
+We measure instantaneous surprisal because we expect the grammaticality of an NPI to have some local effect on the model's predictions.
+We measure cumulative surprisal because we expect the grammaticality of an NPI to have longer-reaching effects on the model's predictions for the whole sentence.
 
@@
 
-As discussed in the introduction, negative polarity items are only grammatically correct in negative contexts.
+As discussed in the introduction, negative polarity items are only grammatically allowable in negative contexts.
 The following examples demonstrate this fact because only the sentences with negative contexts (a,c) are grammatically permissible:
@@
 Plotted in Figure \ref{fig:surprisal-basic-licensing} are graphs of surprisal on two generic examples; negative polarity items are assigned lower surprisal values by the model when they are found in grammatically licit environments.
+
+To determine the statistical significance of the surprisal scores of NPIs in different contexts, we conducted a 1-sample t-test on the interaction score data that we collected from our 20 test sentences.
+The null hypothesis of the test was that the interaction score was not affected by sentence context, and thus our population mean was set to zero.
+However, we found p = .0037, so we reject our null hypothesis and conclude that the upper-tailed alternative hypothesis (that the surprisal with no license is greater than the surprisal with a license) is true.
 

### 2021-12-23 src/main.py
Updates

@@ -209,3 +209,40 @@ def bar():
 batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
 sns.set()
 plot_surprisal(dictionary, model, batch)
+
+
+def read_examples_file_as_tensor_with_licensor(examples, dictionary):
+    """Reads a text file with one example sentence on each line, and returns a
+    tensor of shape (max_len + 1, num_sentences), where max_len is the length of
+    the longest sentence (not including an <eos> token) and num_sentences is the
+    number of sentences."""
+    with open(examples, 'r', encoding="utf8") as f:
+        sequences = []
+        npi_marker_indices = []
+        licensor_indices = []
+        seq_len = 0
+        for line in f:
+            if not line:
+                continue
+
+            words = ['<bos>'] + line.split() + ['<eos>']
+
+            if words[1] == '#':
+                continue
+
+            if '*' not in words or '&' not in words:
+                print("Skipping sentence with missing marker")
+                continue
+            
+            npi_marker_indices.append(words.index('*') - 1)
+            licensor_indices.append(words.index('&'))
+            
+            ids = [word_to_id(dictionary, w) for w in words if w != '*' or w != '&']
+            sequences.append(ids)
+            if len(words) > seq_len:
+                seq_len = len(words)
+
+    eos_id = dictionary.word2idx['<eos>']
+    sequences = [s + [eos_id] * (seq_len - len(s)) for s in sequences]
+    sequences = [torch.tensor(seq).type(torch.int64) for seq in sequences]
+    return torch.stack(sequences), npi_marker_indices, licensor_indices

### 2021-12-23 src/main.py
Add stuff

@@ -193,9 +193,29 @@ def continue_seq(model, dictionary, seq, choices):
 def foo():
     batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
     thing = compute_surprisal_on_batch(model, dictionary, batch, npi_markers)
-    thing = compute_licensing_interaction_paired(thing)
-    thing = [j for i, j in thing]
-    return ttest_1samp(thing, 0, alternative='greater')
+    good = thing[::2]
+    bad = thing[1::2]
+    
+    res = pd.DataFrame()
+    for i, ((good_pnt, good_cum), (bad_pnt, bad_cum)) in enumerate(zip(good, bad)):
+        res = res.append([{'sentence' : i,
+                           'measurement' : 'instant',
+                           'licensed' : 'licensed',
+                           'surprisal' : good_pnt},
+                          {'sentence' : i,
+                           'measurement' : 'instant',
+                           'licensed' : 'unlicensed',
+                           'surprisal' : bad_pnt},
+                          {'sentence' : i,
+                           'measurement' : 'cumulative',
+                           'licensed' : 'licensed',
+                           'surprisal' : good_cum},
+                          {'sentence' : i,
+                           'measurement' : 'cumulative',
+                           'licensed' : 'unlicensed',
+                           'surprisal' : bad_cum}])
+    #    return ttest_1samp(thing, 0, alternative='greater')
+    return res
 
 def bar():
     batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
@@ -205,11 +225,8 @@ def bar():
     thing = pd.Series(thing)
     return thing
     
-    
 batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
 sns.set()
-plot_surprisal(dictionary, model, batch)
-
 
 def read_examples_file_as_tensor_with_licensor(examples, dictionary):
     """Reads a text file with one example sentence on each line, and returns a

### 2021-12-23 paper.tex

@@
 While LSTMs have proven themselves as powerful tools for language modeling tasks, the extent to which their generalizations are based on underlying syntactic structure remains an open area of inquiry.
-We focus our investigation on the relationship between negative polarity items and their licensors in English.
-We probe the model's representation of this relationship by computing the network's surprisal upon encountering NPIs in licensed and unlicensed contexts.
-We show that...
+We focus our investigation on the relationship between negative polarity items (NPIs) and their licensors in English.
+We probe the model's representation of this relationship by computing the network's surprisal upon encountering NPIs in licensed and unlicensed environments.
+While we do observe that the inclusion of a licensor causes a significant reduction in model surprisal, we do not observe any change in surprisal depending on whether or not the licensor c-commands the negative polarity item.
 \end{abstract}
@@
 Optimization was performed using gradient descent, with a base learning rate of 0.2, which decreased exponentially whenever performance on the development set regressed.
+
+The model can be found in the project's GitHub repository (\url{https://github.com/tsandstr/ling380-final-project}).
 
@@
 In this subsection, we confirm the hypothesis that surprisal metrics for NPIs are reduced in the presence of an appropriate licensor.
-We created 20 test items containing negative polarity items.
-For each item, we constructed two variants: one with a licensor, and one without a licensor, as exemplified in Examples \ref{ex:license-1} and \ref{ex:license-2}.
+As described above, we created 20 test items containing negative polarity items, and for each item, we constructed one variant with a licensor, and one without a licensor, as exemplified in Examples \ref{ex:license-1} and \ref{ex:license-2}.
 
 Plotted in Figure \ref{fig:surprisal-basic-licensing} are graphs of surprisal on two generic examples; negative polarity items are assigned lower surprisal values by the model when they are found in grammatically licit environments.
-
-To determine the statistical significance of the surprisal scores of NPIs in different contexts, we conducted a 1-sample t-test on the interaction score data that we collected from our 20 test sentences.
-The null hypothesis of the test was that the interaction score was not affected by sentence context, and thus our population mean was set to zero.
-However, we found p = .0037, so we reject our null hypothesis and conclude that the upper-tailed alternative hypothesis (that the surprisal with no license is greater than the surprisal with a license) is true.
-
-Overall, this experiment shows that our LSTM model has at least a basic understanding of when negative polarity items should be expected.
 
@@
 \end{figure}
+
+To determine the statistical significance of the surprisal scores of NPIs in different contexts, we computed surprisal scores for both variants of our 20 test items and computed the licensing interaction by taking pairwise differences.
+If the model had learned nothing at all about the NPI/licensor relation, we would expect the licensing interaction to be zero.
+However, if the model did learn to predict NPIs only in the presence of a licensor, then the licensing interaction should be positive.
+
+Measuring instantaneous surprisal upon encountering an NPI, we found a mean licensing interaction of 0.87, which we deemed significant ($t(19 = 1.82$, $p = 0.042$), so we reject the null hypothesis, concluding that the sentences containing a licensor had significantly lower suprisal than those without a licensor.
+
+Measuring cumulative surprisal over the remainder of the clause, we found an even larger licensing interaction of 1.83, which we also deemed significant ($t(19) = 3.00$, $p = 0.004$).
+
+Overall, this experiment shows that our LSTM model has at least a basic understanding of when negative polarity items should be expected.
+
+\subsection{C-Command Relation Between Licensor and NPI}
+
+As discussed in the the Introduction, there are structural restrictions on the NPI licensing relationship.
+For an NPI construction to be grammatical, it is not sufficient that a negative licensor precede the NPI in terms of linear word order.
+In fact, the licensor must c-command the NPI in order for the construction to be grammatical.
+
+In the following example, the dependent clause is fixed as negative and the independent clause (which contains the NPI) is varied, so only the first variation is correct:
+\begin{exe}
+\ex
+\begin{xlist}
+\ex Because he was not tired, he was not even sleeping
+\ex [*]{Because he was not tired, he was even sleeping} 
+\end{xlist}
+\end{exe}
+In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items.
+In order to do this we designed 10 base sentences.
+Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
+We are motivated to analyze this type of interaction because a model that is well adjusted to NPI should determine that the presence of negative context in a separate c-command does not influence whether it is actually grammatical.
 
@@
 However as seen in the plots, the scores are relatively similar, so we can conclude that the model does not consider the NPIs as ungrammatical when the amount of intervening words is long.
-\subsection{Licensing of Negative Polarity Items Between Separate C-Commands}
-Just because one section of a sentence is in a negative context does not mean that it is appropriate to use a negative polarity item in another clause.
-In the following example, the dependent clause is fixed as negative and the independent clause (which contains the NPI) is varied, so only the first variation is correct:
-\begin{exe}
-\ex
-\begin{xlist}
-\ex Because he was not tired, he was not even sleeping
-\ex [*]{Because he was not tired, he was even sleeping} 
-\end{xlist}
-\end{exe}
-In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items.
-In order to do this we designed 10 base sentences.
-Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
-We are motivated to analyze this type of interaction because a model that is well adjusted to NPI should determine that the presence of negative context in a separate c-command does not influence whether it is actually grammatical.
+
 \subsection{Ability to License Multiple Negative Polarity Items}

### 2021-12-23 src/main.py
Update graphic

@@ -200,19 +200,19 @@ def foo():
     for i, ((good_pnt, good_cum), (bad_pnt, bad_cum)) in enumerate(zip(good, bad)):
         res = res.append([{'sentence' : i,
                            'measurement' : 'instant',
-                           'licensed' : 'licensed',
+                           'environment' : 'with licensor',
                            'surprisal' : good_pnt},
                           {'sentence' : i,
                            'measurement' : 'instant',
-                           'licensed' : 'unlicensed',
+                           'environment' : 'without licensor',
                            'surprisal' : bad_pnt},
                           {'sentence' : i,
                            'measurement' : 'cumulative',
-                           'licensed' : 'licensed',
+                           'environment' : 'with licensor',
                            'surprisal' : good_cum},
                           {'sentence' : i,
                            'measurement' : 'cumulative',
-                           'licensed' : 'unlicensed',
+                           'environment' : 'without licensor',
                            'surprisal' : bad_cum}])
     #    return ttest_1samp(thing, 0, alternative='greater')
     return res

### 2021-12-23 paper.tex

@@
 \usepackage{caption}
-\usepackage{subfig}
+\usepackage{subcaption}
 \graphicspath{{./img/}}
@@
 For example, in English, negative polarity items include words like \textit{ever}, \textit{any}, and \textit{even}.
-\begin{enumerate}
-    \item He never gave up.
-    \item *He \textit{ever} gave up.
-    \item Nobody ever gave up.
-\end{enumerate}
-Here, the word \textit{ever} is not permitted to occur in a sentence except when licensed by a word such as \textit{nobody} that provides the surrounding "negative" context.
-Notably, the relationship between the licensor and the negative polarity item is based on the hierarchical structure of a sentence, wherein the licensor must c-command the negative polarity item (i.e.
-the negative polarity item must be a sister of the licensor or a direct descendant of a sister node of the licensor).
+\begin{exe}
+\ex \begin{xlist}
+\ex \label{ex:example1} The boy did not ever give up.
+\ex[*]{\label{ex:example2} The boy did ever give up.}
+\ex[*]{\label{ex:example3} The boy who did not win did ever give up.}
+\end{xlist}
+\end{exe}
+Considering Examples \ref{ex:example1} and \ref{ex:example2}, we note that the word \textit{ever} is only permitted to occur in a sentence when it is licensed by a word such as \textit{nobody} which provides a ``negative'' context.
+For this basic example, the relationhip between licensor and NPI would not be a particularly challenging task for a language model to learn; if it doesn't see any negative words in environment, then it won't predict an NPI.
 
-If a language model after being trained on a generic linguistic data set is able to predict the distribution of negative polarity items and the contexts in which they are licensed with significant accuracy, this would be evidence that the model is able to learn this feature of syntax, in particular that it is able to detect a relationship between a licensor and a negative polarity item.
+However, considering Example \ref{ex:example3}, we see that this simple heuristic fails.
+Despite the fact that the word \textit{not} preceeds the NPI, the sentence is still ungrammatical.
+This is because the relationship between the licensor and the negative polarity item is not based on linear word order but rather on the hierarchical structure of a sentence.
+In particular, for an NPI to be permissible, it must be c-commanded by a licensor.
+This c-command relation means that the NPI must be a sister of the licensor or a direct descendant of a sister node of the licensor \cite{ladusaw1980notion}.
+
+If a language model after being trained on a generic linguistic data set is able to predict the distribution of negative polarity items accurately, this would be evidence that the model is able to learn this feature of syntax, and in particular, that it represents hierarchical structure with enough fidelity to capture the c-command relation.
 
@@
 \subsection{Licensing of Negative Polarity Items}
+\label{sec:basic}
 
@@
 However, if the model did learn to predict NPIs only in the presence of a licensor, then the licensing interaction should be positive.
+Our results are plotted in Figure \ref{fig:basic-detection-box-plot}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{surprisal-c-command-box-plot}
+    \caption{Plots of surprisal scores in the presence of a licensor and without a licensor.
+As expected, surprisal tends to be higher in the environment with no licensor.
+Interstingly, \textit{median} instantaneous surprisal actually decreases when we include a licensor.
+However, this does not take into account the fact that our dataset is paired; this is only captured when we apply a $t$-test to the licensing interaction.}
+    \label{fig:basic-detection-box-plot}
+\end{figure}
 
@@
 \subsection{C-Command Relation Between Licensor and NPI}
+\label{sec:c-command}
 
@@
 
-In the following example, the dependent clause is fixed as negative and the independent clause (which contains the NPI) is varied, so only the first variation is correct:
+Consider the sentences in Example \ref{ex:c-command}, each of which contains an embedded clause.
+In \ref{ex:c-command-nn} there is a negative licensor in both the embedded clause and the main clause, in \ref{ex:c-command-np} there is only a licensor in the embedded clause, in \ref{ex:c-command-pn} there is only a licensor in the main clause, and in \ref{ex:c-command-pp} there is no licensor at all.
+
 \begin{exe}
 \ex
+\label{ex:c-command}
 \begin{xlist}
-\ex Because he was not tired, he was not even sleeping
-\ex [*]{Because he was not tired, he was even sleeping} 
+\ex \label{ex:c-command-nn} That car which had no license plate was not even pulled over
+\ex [*]{\label{ex:c-command-np} That car which had no license plate was even pulled over} 
+\ex \label{ex:c-command-pn} That car which had a license plate was not even pulled over
+\ex [*]{\label{ex:c-command-pp} That car which had a license plate was even pulled over}
 \end{xlist}
 \end{exe}
+
 In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items.
-In order to do this we designed 10 base sentences.
-Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
-We are motivated to analyze this type of interaction because a model that is well adjusted to NPI should determine that the presence of negative context in a separate c-command does not influence whether it is actually grammatical.
+In order to do this we designed 15 test items, each containing an embedded clause and an NPI.
+For each item, we generated four variants, as exemplified by those in Example \ref{ex:c-command}.
+If the model has learned that NPIs are only gramatical when they are c-commanded by a licensor, we should expect that the presence of a licensor in the main clause has a significant impact on surprisal, while the presence of a licensor in the embedded clause should not have a significant impact on surpisal.
+The results are visualized in Figure \ref{fig:c-command-box-plots}.
 
-\subsection{Robustness of Negative Polarity Items to Intervening Material}
-Syntactic dependencies are unchanged by the presence intervening material.
-In the following example, the presence of a negative polarity item (anymore) is due to syntactic licensing by negation (don't); modifying the object doesn't change the sentence structure, so it has no effect on NPI licensing:
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{point_surprisal_c_command}
+        \caption{Instantaneous surprisal}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{cummulative_surprisal_c_command}
+        \caption{Cumulative surprisal}
+    \end{subfigure}
+    \caption{Surprisal values for sentences of the form in Example \ref{ex:c-command}.
+The environment of \ref{ex:c-command-nn} corresponds to the left most column, \ref{ex:c-command-np} corresponds to the third column, \ref{ex:c-command-pn} corresponds to the second column, and \ref{ex:c-command-pp} corresponds to the right most column.}
+    \label{fig:c-command-box-plots}
+\end{figure}
+
+To test this hypothesis, we used two-way repeated measures ANOVA to analyze the effect of the embedded licensor and the main licensor on surprisal.
+The test showed a significant main effect of the main licensor on instantaneous surprisal ($F(1, 14) = 7.90$, $p = 0.014$).
+It also showed a significant main effect of the embedded licensor on instantaneous surprisal ($F(1, 14) = 8.22$, $p = 0.012$).
+It did not show any significant interaction between the main licensor and the embedded licensor ($F(1, 14) = 1.87$, $p = 0.193$).
+
+These results tell us two things.
+First, they support the result of section \ref{sec:basic} that the presence of a licensor in the main clause reduces surprisal.
+Second, because there is a significant effect of the embedded licensor, we conclude that the model has not learned to generalize based on underlying sentence structure.
+
+\subsection{Robustness to Intervening Material}
+
+Because the relation between licensor and NPI is based on the c-command relation rather than on linear word order, it is not sensitive to the amount of intervening material which occurs between the licensor and the NPI.
+In Example \ref{ex:intervening}, the NPI \textit{anymore} is licenced by the word \textit{not} in each sentence.
+The intervening prepositional phrases and relative clauses have no impact on the grammaticality of the sentences.
 \begin{exe}
-\ex
+\ex \label{ex:intervening}
 \begin{xlist}
@@
 \end{exe}
-We showed in the previous subsection that LSTMs have expectations for negative polarity items, so we attempt to determine whether those expectations remain over intervening material.
-For this subsection, we designed 10 sentences base sentences with NPI items.
+We showed in Section \ref{sec:basic} that LSTMs exhibit significantly higher surprisal when encountering an unlicensed NPI than a licensed NPI; however, the analysis in Section \ref{sec:c-command} suggests that the model is using a heuristic based on linear word order rather than taking into account the hierarchical structure of the sentence.
+Given this, we might predict that the amount of intervening material could have an impact on surprisal.
+
+To test the robustness of the model's predictions to intervening material, we first designed 21 test items containing NPIs.
+In constructing these items, we made sure to vary the amount of intervening material between the licensor and the NPI.
+Then, for each test item, we constructed two variations: one with a licensor and one without.
+
+We also did a slightly different experiment aiming to control for variation in the semantic content of the sentences.
+Here, we designed 10 sentences base sentences with NPI items.
 Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item.
 In all cases the intervening words modify the object of the base sentence.
+An example of the results for this section are visualized in Figure 2.
 
@@
 
-An example of the results for this section are visualized in Figure 2.
-We found that in general, there was a weak positive correlation between the number of words in between the word that determined the negative context and the negative polarity item and the surprisal score.
-However as seen in the plots, the scores are relatively similar, so we can conclude that the model does not consider the NPIs as ungrammatical when the amount of intervening words is long.
+We once again found that there was a weak positive correlation between amount of intervening material and the licensing interaction.
+However as seen in the plots, the scores are relatively similar.
+This test supports the notation that the model's ability to predict the the distribution of NPIs is not affected by the quantity of intervening material.
 

### 2021-12-24 src/main.py
Update

@@ -217,11 +217,11 @@ def foo():
     #    return ttest_1samp(thing, 0, alternative='greater')
     return res
 
-def bar():
-    batch, npi_markers = read_examples_file_as_tensor(args.evaluate, dictionary)
+def bar(filename = args.evaluate):
+    batch, npi_markers = read_examples_file_as_tensor(filename, dictionary)
     thing = compute_surprisal_on_batch(model, dictionary, batch, npi_markers)
     thing = compute_licensing_interaction_paired(thing)
-    thing = [j for i, j in thing]
+    thing = [i for i, j in thing]
     thing = pd.Series(thing)
     return thing
     

### 2021-12-24 src/main.py
Rearrange files

@@ -14,7 +14,7 @@ import model
 parser = argparse.ArgumentParser(description='Evaluate language model on examples')
 parser.add_argument('--model', type=str, default='lstm.pt',
                     help="location of model to evaluate")
-parser.add_argument('--evaluate', type=str, default='../sentences.txt',
+parser.add_argument('--evaluate', type=str, default='../experiment_data/sentences.txt',
                     help='location of example sentences to evaluate')
 parser.add_argument('--dictionary', type=str, default='dictionary.p',
                     help='location of the dictionary to use')

### 2021-12-24 paper/paper.tex

--- 2021-12-23 paper.tex
@@
 There is, though, little research investigating the feature of licensing of negative polarity items in language.
-Negative polarity items are a class of words that are only permitted to appear in certain "negative" contexts.
+Negative polarity items are a class of words that are only permitted to appear in certain ``negative'' contexts.
 For example, in English, negative polarity items include words like \textit{ever}, \textit{any}, and \textit{even}.
@@
 
-If a language model after being trained on a generic linguistic data set is able to predict the distribution of negative polarity items accurately, this would be evidence that the model is able to learn this feature of syntax, and in particular, that it represents hierarchical structure with enough fidelity to capture the c-command relation.
+If a language model trained on a generic linguistic data set is able to predict the distribution of negative polarity items accurately, this would be evidence that the model is able to learn this feature of syntax, and in particular, that its internal representation captures hierarchical structure with enough fidelity to encapsulate the c-command relation.
 
@@
 
-For our investigation, we trained a LSTM model on the WikiText-2 dataset \cite{merity2016pointer}, which consists of over 2 million tokens from 600 Wikipedia articles, with a vocabulary size of 33,279.
+For our investigation, we trained a LSTM model on the WikiText-2 data set \cite{merity2016pointer}, which consists of over 2 million tokens from 600 Wikipedia articles, with a vocabulary size of 33,279.
 Our LSTM has an embedding layer, assigning 200-dimensional embeddings to each word in the vocabulary, and two recurrent layers, each with 650 hidden units.
@@
 
-The model can be found in the project's GitHub repository (\url{https://github.com/tsandstr/ling380-final-project}).
+The model can be found in the project's GitHub repository\footnote{\url{https://github.com/tsandstr/ling380-final-project}}.
 
@@
 A high surprisal metric means that, in the given context, a word was predicted with \textit{low} probability, and was therefore unexpected by the model.
-Basically, given a hidden state of the model at a current time step, the surprisal metric is a way to measure the whether the next word makes \textit{sense}.
+Intuitively, given a hidden state of the model at a current time step, the surprisal metric is a way to measure the whether the next word makes \textit{sense}.
 A direct correlation has been shown between language model surprisal metrics and  human sentence processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-
-\subsection{Experimental Design}
 
@@
 Another property is that NPIs are immune to intervening material separating the NPI from its licensor.
-And finally, one negative context can license multiple negative polarity items.
-In this section we demonstrate that LSTMs have learned these four properties negative polarity items.
+In this section we  investigate whether LSTMs understand these three properties.
 
@@
 Interstingly, \textit{median} instantaneous surprisal actually decreases when we include a licensor.
-However, this does not take into account the fact that our dataset is paired; this is only captured when we apply a $t$-test to the licensing interaction.}
+However, this does not take into account the fact that our data set is paired; this is only captured when we apply a $t$-test to the licensing interaction.}
     \label{fig:basic-detection-box-plot}
@@
 
+Comparing the two statistically significant effects, we used partial-$\eta^2$ to measure the size of these effects (the (partial)-$\eta^2$ statistic for ANOVA is analogous to the $r^2$ statistic for linear regression, in that it described the proportion of variance in the dependent variable explained by each independent variable).
+We found that the effect size for the main licensor is $\eta_p^2 = 0.36$ and the effect size for the embedded licensor is $\eta_p^2 = 0.37$.
+Thus, both the main licensor and the embedded licensor have similar sized effets on instantaneous surprisal.
+
+The fact that the interaction term is not significant means that the effect of the embedded licensor does not depend on the presence or lack of a main licensor.
+In other words, the effects of the licensors combine linearly.
+
 These results tell us two things.
@@
 \subsection{Robustness to Intervening Material}
-
+\label{sec:intervening}
 Because the relation between licensor and NPI is based on the c-command relation rather than on linear word order, it is not sensitive to the amount of intervening material which occurs between the licensor and the NPI.
@@
 Then, for each test item, we constructed two variations: one with a licensor and one without.
+For each item, we computed the licensing interaction, and then checked for a correlation between the amount of intervening material and the licensing interaction.
+Our results are visualized in Figure \ref{fig:intervening1}.
+We found no statistically signficant correlation between intervening length and either instantaneous licensing interaction ($p = 0.305$) or cumulative licensing interaction ($p = 0.485$).
+This suggests that the model's ability to predict NPI distribution is not affected by the mount of intervening material between the NPI and its licensor.
 
-We also did a slightly different experiment aiming to control for variation in the semantic content of the sentences.
-Here, we designed 10 sentences base sentences with NPI items.
-Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item.
-In all cases the intervening words modify the object of the base sentence.
-An example of the results for this section are visualized in Figure 2.
-
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{intervening_pnt_scatter}
+        \caption{Instantaneous licensing interaction}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{intervening_cum_scatter}
+        \caption{Cumulative licensing interaction}
+    \end{subfigure}
+    \caption{Plots of licensing interaction of sentences given the amount of words in between the negative licensor and the NPI}
+    \label{fig:intervening1}
+\end{figure}
 \begin{figure*}
@@
 The surprisal score of the negative polarity item \textit{anymore} remains relatively unchanged by the amount of intervening material.}
+    \label{fig:intervening2}
 \end{figure*}
+We also did a slightly different experiment aiming to control for variation in the semantic content of the sentences.
+Here, we designed 10 sentences base sentences with NPI items.
+Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item.
+In all cases the intervening words modify the object of the base sentence.
+An example of the results for this section are visualized in Figure \ref{fig:intervening2}.
 
 We once again found that there was a weak positive correlation between amount of intervening material and the licensing interaction.
-However as seen in the plots, the scores are relatively similar.
+However, as seen in the plots, the scores are relatively similar.
 This test supports the notation that the model's ability to predict the the distribution of NPIs is not affected by the quantity of intervening material.
 
-\subsection{Ability to License Multiple Negative Polarity Items}
-In most cases, a negative context can license multiple negative polarity items.
+% \subsection{Ability to License Multiple Negative Polarity Items}
+% In most cases, a negative context can license multiple negative polarity items.
+% \begin{exe}
+% \ex
+% \begin{xlist}
+% \ex He didn't want to go with anybody anywhere
+% \ex [*]{He did want to go with anybody anywhere} 
+% \end{xlist}
+% \end{exe} 
+\section{Positive Polarity Items}
+We briefly turn our attention to positive polarity items (PPIs) which are words that can only occur in ``positive'' contexts.
+English words such as \textit{some, rather,} and \textit{already} are common examples of PPIs.
+There is significantly less literature discussing PPIs than NPIs, and often the licensing of a positive context is much less explicit than that of a negative context.
+We are interested in whether LSTMs will exhibit higher surprisal in unlicensed settings, and if so, does the model understand PPIs to the same extent as NPIs.
+Establishing a positive context in English is not as explicit as establishing a negative context.
+Sentences default to the positive setting, and thus there is not an equivalent to \textit{not} which licenses a negative context.
+We hypothesize that the lack of an obvious licensor might make it more difficult for the model to determine whether a PPI is permissible.
+
+\subsection{Licensing of Positive Polarity Items}
+Positive polarity items may only be licensed in positive contexts, as is shown for the PPI \textit{rather} in the following example.
 \begin{exe}
-\ex
+\ex \label{ex:positive}
 \begin{xlist}
-\ex He didn't want to go with anybody anywhere
-\ex [*]{He did want to go with anybody anywhere} 
+\ex I would rather be there
+\ex[*]{I wouldn't rather be there}
+
 \end{xlist}
 \end{exe}
+Plotted in Figure \ref{fig:ppi-basic-licensing} are graphs surprisal of the sentences in \ref{ex:positive}.
+One can see that the positive polarity item has a slightly lower surprisal score in the grammatically correct context, but the trend is not as pronounced as what was seen with NPIs.
 
-\section{Results}
+We conduct a nearly identical experiment as that in Section \ref{sec:basic} to determine the extent to which LSTMs learn about PPIs.
+We constructed 10 test items containing positive polarity items, and for each item, we constructed one with a positive context, and another with a negative context as exemplified in Example \ref{ex:positive}.
+If the model learned to predict PPIs only in positive contexts, then we would expect the licensing interaction of these sentences to be positive.
+
+\begin{figure}
+    \centering
+    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{img/ppigood.png} }}
+    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{img/ppibad.png} }}
+     \caption{The surprisal score for \textit{rather} is slightly lower in the positive context than in the incorrect negative context}
+    \label{fig:ppi-basic-licensing}
+\end{figure}
+
+Measuring both instantaneous surprisal and cumulative surprisal after a PPI, we found in both cases that the mean licensing interaction is slightly greater than 0 (which implies that surprisal is higher when we find a PPI in a negative context), but the difference is not large enough to be statistically significant (instantaneous, $t(9) = 1.43$, $p = 0.093$; cumulative, $t(9) = 1.37$, $p = 0.102$).
+Thus, we fail to reject the null hypothesis that the licensing interaction is zero.
+
+This small experiment suggests that it is more difficult for an LSTM to determine the validity of a PPI than an NPI.
+This result is not particularly surprising because, unlike NPIs, PPIs lack explicit licensors.
+Thus, in order to correctly predict the distribution of PPIs, the model would need to represent more complex aspects of language.
+
+\section{Results and Discussion}
+
+In Section \ref{sec:basic}, we showed that the presence of a negative licensor in a sentence caused a significant decrease in the model's surprisal upon encountering an NPI.
+This suggests that the model has learned that NPIs only occur when a negative licensor is present.
+However, in Section \ref{sec:c-command}, we saw that the model does not distinguish between licensors which c-command the NPI and those that do not.
+This suggests that, rather than representing the hierarchical structure of the sentence, the model approximates the NPI/licensor relation using a representation based on linear word order and precedence.
+Finally, in Section \ref{sec:intervening}, we saw that the model's ability to predict the distribution of NPIs was independent of the amount of intervening material.
+
+Despite the fact that the experiments in Section \ref{sec:c-command} suggested that the model does not represent the c-command relation, the results do not differ too dramatically from what is observed in psycholinguistic studies of NPI licensing.
+As we would expect, listeners with no time constraints tend to judge sentences like those in Examples \ref{ex:c-command-np} and \ref{ex:c-command-pp} as being ungrammatical.
+However, when fast time constraints are placed on the judgement task, some speakers mistakenly accept sentences like Example \ref{ex:c-command-np} (in which there exists a negative licensor, but it is contained in an embedded clause) as grammatical \cite{xiang2009illusory,xiang2013dependency,parker2016negative}.
+This suggests that, under time pressure, the human brain might process the NPI licensing relation using a heuristic similar to the one that our model has learned, which takes into account only precedence and not c-command.
+
+These results point towards a number of questions which could be addressed in future research.
+First of all, we might like to know how well these results generalize to other models.
+Our homegrown model was trained on a relatively small data set using limited resources; perhaps a larger LSTM like that presented in Jozefowicz et.
+al.
+\shortcite{jozefowicz2016exploring} would have a better chance of correctly modeling the c-command relationship.
+We could also apply a similar analysis to other types of network architectures, such as transformers.
+Large transformers like the GPT-2 model presented by Radford et.
+al.
+\shortcite{radford2019language} and the BERT model introduced in Devlin et.
+al.
+\shortcite{devlin2018bert} have both performed very well on a very general set of language modeling tasks.
+Repeating a similar analysis on GPT-2 or Bert could provide interesting insights into the extent to which transformer models represent hierarchical structure.
+
+Of course, there is also further research that could be done without moving to a new model or architecture.
+One possibility would be to try to identify a specific unit or collection of units in the cell state of the LSTM that is responsible for tracking negative and positive states.
+This sort of analysis has been achieved before; in Lakretz et.
+al.
+\shortcite{lakretz2019emergence}, the authors identified discrete units responsible for tracking information about number agreement.
+They achieved this by using a technique known as \textit{ablation}.
+First, they constructed a task which measured the models ability to track number agreement.
+Then, one by one, they zeroed out (ablated) individual units in the LSTMs cell state.
+By selecting units whose ablation was associated with decreased success on the number agreement task, they were able to identify specific units responsible for tracking number agreement.
+
+If a similar ablative analysis could identify specific units responsible for tracking negative polarity licensing, we could make predictions about how this unit's activation would change over the course of a sentence.
+In Section \ref{sec:c-command}, we observed that both embedded licensors and main licensors have a significant effect on surprisal, and that these effects combine linearly.
+This suggests that a hypothetical ``NPI licensing unit'' in the LSTM's hidden state should be expected to increase in activation each time the model encounters a negative word, regardless of the current level of activation.
+If we could observe this pattern in a specific unit of the LSTM's cell state, this would confirm that the unit identified by the ablative analysis was in fact responsible for tracking NPI licensing.
 

### 2021-12-24 paper.tex
updates

@@
 Historically, the majority of research on these networks has focused simply on evaluating and optimizing performance on applied NLP tasks.
-However, there has been a recent increase in literature probing inside of these black box models and evaluating the extent to which their success is based on surface level heuristics rather than genuinely extracting and interpreting structural properties of language, either syntactic or semantic.
-\cite{lakretz2019emergence}.
+However, there has been a recent increase in literature probing inside of these black box models and evaluating the extent to which their success is based on surface level heuristics rather than genuinely extracting and interpreting structural properties of language, either syntactic or semantic \cite{lakretz2019emergence}.
 
@@
 
-If a language model after being trained on a generic linguistic data set is able to predict the distribution of negative polarity items accurately, this would be evidence that the model is able to learn this feature of syntax, and in particular, that it represents hierarchical structure with enough fidelity to capture the c-command relation.
+Jumelet and Hupkes \shortcite{jumelet2018language} investigated neural language models for NPIs via analysis utilizing perplexity scores, defined as the exponent of the normalized negative log probability of different sentences, wherein the lower the perplexity score of a sentence is, the better a model was able to predict its tokens.
+Jumelet et al.
+\shortcite{jumelet2021language} similarly investigated neural language models for NPIs via analysis utilizing conditional probability distributions and a novel method ranking their model's decoder weights for all tokens based on their similarity with monotonicity weights in their sentence data sets.
+This is the general extent of the existing literature on NPIs, and neither study makes reference to the surprisal metric, which we introduce in Subsection \ref{sec:surpisal} and contend is a superior metric for analysis.
+
+If a language model trained on a generic linguistic data set is able to predict the distribution of negative polarity items accurately, this would be evidence that the model is able to learn this feature of syntax, and in particular, that its internal representation captures hierarchical structure with enough fidelity to encapsulate the c-command relation.
 
@@
 
-The model can be found in the project's GitHub repository (\url{https://github.com/tsandstr/ling380-final-project}).
+The model can be found in the project's GitHub repository\footnote{\url{https://github.com/tsandstr/ling380-final-project}}.
 
 \subsection{Surprisal}
+\label{sec:surpisal}
 
@@
 The surprisal score of the negative polarity item \textit{anymore} remains relatively unchanged by the amount of intervening material.}
+    \label{fig:intervening2}
 \end{figure*}
+We also did a slightly different experiment aiming to control for variation in the semantic content of the sentences.
+Here, we designed 10 sentences base sentences with NPI items.
+Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item.
+In all cases the intervening words modify the object of the base sentence.
+An example of the results for this section are visualized in Figure \ref{fig:intervening2}.
 
 We once again found that there was a weak positive correlation between amount of intervening material and the licensing interaction.
-However as seen in the plots, the scores are relatively similar.
+However, as seen in the plots, the scores are relatively similar.
 This test supports the notation that the model's ability to predict the the distribution of NPIs is not affected by the quantity of intervening material.
 
-\subsection{Ability to License Multiple Negative Polarity Items}
-In most cases, a negative context can license multiple negative polarity items.
+% \subsection{Ability to License Multiple Negative Polarity Items}
+% In most cases, a negative context can license multiple negative polarity items.
+% \begin{exe}
+% \ex
+% \begin{xlist}
+% \ex He didn't want to go with anybody anywhere
+% \ex [*]{He did want to go with anybody anywhere} 
+% \end{xlist}
+% \end{exe} 
+\section{Positive Polarity Items}
+We briefly turn our attention to positive polarity items (PPIs) which are words that can only occur in ``positive'' contexts.
+English words such as \textit{some, rather,} and \textit{already} are common examples of PPIs.
+There is significantly less literature discussing PPIs than NPIs, and often the licensing of a positive context is much less explicit than that of a negative context.
+We are interested in whether LSTMs will exhibit higher surprisal in unlicensed settings, and if so, does the model understand PPIs to the same extent as NPIs.
+Establishing a positive context in English is not as explicit as establishing a negative context.
+Sentences default to the positive setting, and thus there is not an equivalent to \textit{not} which licenses a negative context.
+We hypothesize that the lack of an obvious licensor might make it more difficult for the model to determine whether a PPI is permissible.
+
+\subsection{Licensing of Positive Polarity Items}
+Positive polarity items may only be licensed in positive contexts, as is shown for the PPI \textit{rather} in the following example.
 \begin{exe}
-\ex
+\ex \label{ex:positive}
 \begin{xlist}
-\ex He didn't want to go with anybody anywhere
-\ex [*]{He did want to go with anybody anywhere} 
+\ex I would rather be there
+\ex[*]{I wouldn't rather be there}
+
 \end{xlist}
 \end{exe}
+Plotted in Figure \ref{fig:ppi-basic-licensing} are graphs surprisal of the sentences in \ref{ex:positive}.
+One can see that the positive polarity item has a slightly lower surprisal score in the grammatically correct context, but the trend is not as pronounced as what was seen with NPIs.
 
-\section{Results}
+We conduct a nearly identical experiment as that in Section \ref{sec:basic} to determine the extent to which LSTMs learn about PPIs.
+We constructed 10 test items containing positive polarity items, and for each item, we constructed one with a positive context, and another with a negative context as exemplified in Example \ref{ex:positive}.
+If the model learned to predict PPIs only in positive contexts, then we would expect the licensing interaction of these sentences to be positive.
+
+\begin{figure}
+    \centering
+    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{img/ppigood.png} }}
+    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{img/ppibad.png} }}
+     \caption{The surprisal score for \textit{rather} is slightly lower in the positive context than in the incorrect negative context}
+    \label{fig:ppi-basic-licensing}
+\end{figure}
+
+Measuring both instantaneous surprisal and cumulative surprisal after a PPI, we found in both cases that the mean licensing interaction is slightly greater than 0 (which implies that surprisal is higher when we find a PPI in a negative context), but the difference is not large enough to be statistically significant (instantaneous, $t(9) = 1.43$, $p = 0.093$; cumulative, $t(9) = 1.37$, $p = 0.102$).
+Thus, we fail to reject the null hypothesis that the licensing interaction is zero.
+
+This small experiment suggests that it is more difficult for an LSTM to determine the validity of a PPI than an NPI.
+This result is not particularly surprising because, unlike NPIs, PPIs lack explicit licensors.
+Thus, in order to correctly predict the distribution of PPIs, the model would need to represent more complex aspects of language.
+
+\section{Results and Discussion}
+
+In Section \ref{sec:basic}, we showed that the presence of a negative licensor in a sentence caused a significant decrease in the model's surprisal upon encountering an NPI.
+This suggests that the model has learned that NPIs only occur when a negative licensor is present.
+However, in Section \ref{sec:c-command}, we saw that the model does not distinguish between licensors which c-command the NPI and those that do not.
+This suggests that, rather than representing the hierarchical structure of the sentence, the model approximates the NPI/licensor relation using a representation based on linear word order and precedence.
+Finally, in Section \ref{sec:intervening}, we saw that the model's ability to predict the distribution of NPIs was independent of the amount of intervening material.
+
+Despite the fact that the experiments in Section \ref{sec:c-command} suggested that the model does not represent the c-command relation, the results do not differ too dramatically from what is observed in psycholinguistic studies of NPI licensing.
+As we would expect, listeners with no time constraints tend to judge sentences like those in Examples \ref{ex:c-command-np} and \ref{ex:c-command-pp} as being ungrammatical.
+However, when fast time constraints are placed on the judgement task, some speakers mistakenly accept sentences like Example \ref{ex:c-command-np} (in which there exists a negative licensor, but it is contained in an embedded clause) as grammatical \cite{xiang2009illusory,xiang2013dependency,parker2016negative}.
+This suggests that, under time pressure, the human brain might process the NPI licensing relation using a heuristic similar to the one that our model has learned, which takes into account only precedence and not c-command.
+
+These results point towards a number of questions which could be addressed in future research.
+First of all, we might like to know how well these results generalize to other models.
+Our homegrown model was trained on a relatively small data set using limited resources; perhaps a larger LSTM like that presented in Jozefowicz et.
+al.
+\shortcite{jozefowicz2016exploring} would have a better chance of correctly modeling the c-command relationship.
+We could also apply a similar analysis to other types of network architectures, such as transformers.
+Large transformers like the GPT-2 model presented by Radford et.
+al.
+\shortcite{radford2019language} and the BERT model introduced in Devlin et.
+al.
+\shortcite{devlin2018bert} have both performed very well on a very general set of language modeling tasks.
+Repeating a similar analysis on GPT-2 or Bert could provide interesting insights into the extent to which transformer models represent hierarchical structure.
+
+Of course, there is also further research that could be done without moving to a new model or architecture.
+One possibility would be to try to identify a specific unit or collection of units in the cell state of the LSTM that is responsible for tracking negative and positive states.
+This sort of analysis has been achieved before; in Lakretz et.
+al.
+\shortcite{lakretz2019emergence}, the authors identified discrete units responsible for tracking information about number agreement.
+They achieved this by using a technique known as \textit{ablation}.
+First, they constructed a task which measured the models ability to track number agreement.
+Then, one by one, they zeroed out (ablated) individual units in the LSTMs cell state.
+By selecting units whose ablation was associated with decreased success on the number agreement task, they were able to identify specific units responsible for tracking number agreement.
+
+If a similar ablative analysis could identify specific units responsible for tracking negative polarity licensing, we could make predictions about how this unit's activation would change over the course of a sentence.
+In Section \ref{sec:c-command}, we observed that both embedded licensors and main licensors have a significant effect on surprisal, and that these effects combine linearly.
+This suggests that a hypothetical ``NPI licensing unit'' in the LSTM's hidden state should be expected to increase in activation each time the model encounters a negative word, regardless of the current level of activation.
+If we could observe this pattern in a specific unit of the LSTM's cell state, this would confirm that the unit identified by the ablative analysis was in fact responsible for tracking NPI licensing.
+
+\section{Further Areas of Inquiry}
+
+The processing of sentences producing Negative Polarity Illusions remains a remarkably interesting open area of inquiry.
+Such illusions refer to sentences wherein there is an illusory licensing of agreement and negative polarity items, in that sentences with unlicensed agreement or an unlicensed NPI are perceived as comprehensible and acceptable upon first utterance but upon further reflection are judged as unacceptable due to faulty licensing.
+Analyzing if LSTM models are able to distinguish in instinctive licensing "correctness," wherein sentences non-illusory sentences are perceived as acceptable or unacceptable strictly contingent upon licensing and illusory sentences are perceived as acceptable or unacceptable in a more loosely fashion, is a goal for further research.
 

### 2021-12-24 paper/paper.tex
Updates

@@
 \usepackage{latexsym}
+
+\usepackage[linguistics]{forest}
 
@@
 \end{exe}
-Considering Examples \ref{ex:example1} and \ref{ex:example2}, we note that the word \textit{ever} is only permitted to occur in a sentence when it is licensed by a word such as \textit{nobody} which provides a ``negative'' context.
-For this basic example, the relationhip between licensor and NPI would not be a particularly challenging task for a language model to learn; if it doesn't see any negative words in environment, then it won't predict an NPI.
+Considering Examples \ref{ex:example1} and \ref{ex:example2}, we note that the word \textit{ever} is only permitted to occur in a sentence when it is licensed by a word such as \textit{not} which provides a ``negative'' context.
+For this basic example, the relationship between licensor and NPI would not be a particularly challenging task for a language model to learn; if it doesn't see any negative words in environment, then it won't predict an NPI.
 
@@
 This c-command relation means that the NPI must be a sister of the licensor or a direct descendant of a sister node of the licensor \cite{ladusaw1980notion}.
+Considering the tree below, it is clear that the licensor \textit{not} fails to c-command the NPI \textit{ever}, so the sentence is ungrammatical.
+
+\begin{exe}
+\ex \begin{forest}
+[TP [NP [The] [N' [N [boy]] [CP [C [who]] [TP [T [did] [not,circle,fill=lightgray]] [VP [win]]]]]] [T' [T [did] [ever,draw,red]] [VP [give up]]]]
+\end{forest}
+\end{exe}
 
 If a language model trained on a generic linguistic data set is able to predict the distribution of negative polarity items accurately, this would be evidence that the model is able to learn this feature of syntax, and in particular, that its internal representation captures hierarchical structure with enough fidelity to encapsulate the c-command relation.
+
+Jumelet and Hupkes \shortcite{jumelet2018language} investigated neural language models for NPIs via analysis utilizing the \textit{perplexity} metric.
+This metric is a measure of how well a language model predicts a certain test item; the lower the perplexity score of a sentence is, the better a model was able to predict its tokens.
+Jumelet et al.
+\shortcite{jumelet2021language} similarly investigated neural language models for NPIs via analysis utilizing conditional probability distributions and a novel method ranking their model's decoder weights for all tokens based on their similarity with monotonicity weights in their sentence data sets.
+This is the general extent of the existing literature on NPIs, and neither study makes reference to the surprisal metric, which we introduce in Subsection \ref{sec:surpisal} and contend is a superior metric for analysis.
 
@@
 \subsection{Surprisal}
+\label{sec:surpisal}
 
@@
 \[ S(x_i) = -\log_2 P(x_i \, | \, h_{i-1}), \]
-where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is given by the softmax activation.
+where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state prior to $x_i$ and the probability is given by the softmax activation.
 The surprisal is measured in bits because we take a base two logarithm.
@@
 A direct correlation has been shown between language model surprisal metrics and  human sentence processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+
+Surprisal is closely related to the perplexity metric used by Jumelet et.
+al.
+\shortcite{jumelet2018language} to evaluate an LSTM's ability to predict NPI distribution.
+While surpisal can be thought of as a measure of how unexpected a given word is in a sentence, perplexity measures how unexpected the entire sentence is.
+By measuring surprisal on specific words, or on specific parts of sentences, we are able to achieve a finer-grained analysis than would be possible using only perplexity.
 
@@
 \[ \textrm{interaction} = S_{\textrm{no license}} - S_{\textrm{license}}. \]
-A large licensing interaction suggests that, in a given test item, the presence of a licensor does dampen the model's surprisal at encountering an NPI, while a small licensing interactoin suggests that, in a given test item, the licensor does not dampen the surprisal.
+A positive licensing interaction suggests that, in a given test item, the presence of a licensor does suppress the model's surprisal upon encountering an NPI, while a zero (or negative) licensing interaction suggests that, in a given test item, the licensor does not dampen the surprisal.
+Thus, if LSTMs are able to correctly predict the grammatically of an NPI based on the presence, or lack thereof, of a negative context we would expect to see a statistically significant licencing interaction.
 
@@
 
-As discussed in the introduction, negative polarity items are only grammatically allowable in negative contexts.
-The following examples demonstrate this fact because only the sentences with negative contexts (a,c) are grammatically permissible:
+As discussed in the introduction, negative polarity items are only permissible in negative contexts.
+The following examples demonstrate this fact because only the sentences with negative contexts are grammatically permissible:
 \begin{exe}
@@
     \centering
-    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{good1} }}
-    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad1} }}
+    \subfloat[Grammatical]{{\includegraphics[width=.9\linewidth]{good1} }}
     \qquad
-    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{good2} }}
-    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad2} }}
+    \subfloat[Ungrammatical]{{\includegraphics[width=.9\linewidth]{bad1} }}
+    \qquad
+    \subfloat[Grammatical]{{\includegraphics[width=.9\linewidth]{good2} }}
+    \qquad
+    \subfloat[Ungrammatical]{{\includegraphics[width=.9\linewidth]{bad2} }}
     \caption{Plots of surprisal scores for words in the example sentences.
@@
 
-Measuring instantaneous surprisal upon encountering an NPI, we found a mean licensing interaction of 0.87, which we deemed significant ($t(19 = 1.82$, $p = 0.042$), so we reject the null hypothesis, concluding that the sentences containing a licensor had significantly lower suprisal than those without a licensor.
+Measuring instantaneous surprisal upon encountering an NPI, we found a mean licensing interaction of 0.87, which we deemed significant ($t(19) = 1.82$, $p = 0.042$), so we reject the null hypothesis that the instantaneous surprisal of a negative polarity item is unchanged by the sentence context.
+Therefore, we conclude that sentences containing a negative licensor had significantly lower suprisal than those without a licensor.
 
@@
 
-Overall, this experiment shows that our LSTM model has at least a basic understanding of when negative polarity items should be expected.
+We conclude that LSTMs understand that the presence of a negative licensor in a sentence increases the probability that a negative polarity item will be present.
+Overall, this signifies that our LSTM model has at least a basic understanding of when negative polarity items should be expected.
 
@@
 
-As discussed in the the Introduction, there are structural restrictions on the NPI licensing relationship.
+As discussed in the Introduction, there are structural restrictions on the NPI licensing relationship.
 For an NPI construction to be grammatical, it is not sufficient that a negative licensor precede the NPI in terms of linear word order.
@@
 
-In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items.
-In order to do this we designed 15 test items, each containing an embedded clause and an NPI.
+In order to analyze the model's performance on sentences like these, we designed 15 test items, each containing an embedded clause and an NPI.
 For each item, we generated four variants, as exemplified by those in Example \ref{ex:c-command}.
@@
 These results tell us two things.
-First, they support the result of section \ref{sec:basic} that the presence of a licensor in the main clause reduces surprisal.
+First, they support the result of Section \ref{sec:basic} that the presence of a licensor in the main clause reduces surprisal.
 Second, because there is a significant effect of the embedded licensor, we conclude that the model has not learned to generalize based on underlying sentence structure.
+Instead, it appears that the model incorrectly predicts that the presence of a negative licensor, regardless of its position in a sentence, increases the probability for a negative polarity item to exist.
 
@@
 We showed in Section \ref{sec:basic} that LSTMs exhibit significantly higher surprisal when encountering an unlicensed NPI than a licensed NPI; however, the analysis in Section \ref{sec:c-command} suggests that the model is using a heuristic based on linear word order rather than taking into account the hierarchical structure of the sentence.
-Given this, we might predict that the amount of intervening material could have an impact on surprisal.
+Given this, we might predict that the amount of intervening material could have an impact on surprisal because the model could forget about the previous presence of a negative licensor as more words are placed between it and an NPI.
 
@@
     \centering
-    \subfloat[Base Sentence]{{\includegraphics[width=.32\textwidth]{short} }}
-    \subfloat[5 Additional Words]{{\includegraphics[width=.32\textwidth]{medium} }}
-    \subfloat[8 Additional Words]{{\includegraphics[width=.32\textwidth]{long} }}
+    \subfloat[Base Sentence]{{\includegraphics[width=.49\textwidth]{short} }}
+    \subfloat[8 Additional Words]{{\includegraphics[width=.49\textwidth]{long} }}
     \caption{Plots of surprisal scores for words in the modified sentence.
-The surprisal score of the negative polarity item \textit{anymore} remains relatively unchanged by the amount of intervening material.}
+The surprisal score of the negative polarity item \textit{anymore} remains relatively unchanged by the amount of intervening material.
+We see that the instantaneous surprisal of the NPI  differs by under 14\% in this example.}
     \label{fig:intervening2}
@@
 An example of the results for this section are visualized in Figure \ref{fig:intervening2}.
+One can see that the length of intervening material has a marginal effect on the surprisal metric on the negative polarity item.
 
 We once again found that there was a weak positive correlation between amount of intervening material and the licensing interaction.
-However, as seen in the plots, the scores are relatively similar.
-This test supports the notation that the model's ability to predict the the distribution of NPIs is not affected by the quantity of intervening material.
+However as seen in Figure \ref{fig:intervening2}, which are a good representation of the overall data, the scores are relatively similar.
+This test supports the notion that the model's ability to predict the the distribution of NPIs is not significantly affected by the quantity of intervening material.
 
@@
 \section{Positive Polarity Items}
-We briefly turn our attention to positive polarity items (PPIs) which are words that can only occur in ``positive'' contexts.
+We briefly turn our attention to another class of words that are off interest, positive polarity items (PPIs).
+PPIs can only occur in ``positive'' contexts.
 English words such as \textit{some, rather,} and \textit{already} are common examples of PPIs.
@@
 Establishing a positive context in English is not as explicit as establishing a negative context.
-Sentences default to the positive setting, and thus there is not an equivalent to \textit{not} which licenses a negative context.
+Declarative sentences in English default to a positive setting, and thus there is not an equivalent to \textit{not} which licenses a negative context.
 We hypothesize that the lack of an obvious licensor might make it more difficult for the model to determine whether a PPI is permissible.
@@
 \subsection{Licensing of Positive Polarity Items}
-Positive polarity items may only be licensed in positive contexts, as is shown for the PPI \textit{rather} in the following example.
+\label{sec:ppi}
+Positive polarity items may only be licensed in positive contexts, as is shown for the PPIs \textit{rather} and \textit{already} in the following examples.
 \begin{exe}
@@
 \ex[*]{I wouldn't rather be there} 
+\end{xlist}
+\ex \label{ex:positive2}
+\begin{xlist}
+\ex Evan is already in jail
+\ex[*]{Evan isn't already in jail}
 
@@
 One can see that the positive polarity item has a slightly lower surprisal score in the grammatically correct context, but the trend is not as pronounced as what was seen with NPIs.
+These plots are a good representation of the general trend that we observed with positive polarity items in different sentence contexts.
 
@@
     \centering
-    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{img/ppigood.png} }}
-    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{img/ppibad.png} }}
+    \subfloat[Grammatical]{{\includegraphics[width=.8\linewidth]{img/ppigood.png} }}
+    \qquad
+    \subfloat[Ungrammatical]{{\includegraphics[width=.8\linewidth]{img/ppibad.png} }}
      \caption{The surprisal score for \textit{rather} is slightly lower in the positive context than in the incorrect negative context}
@@
 This result is not particularly surprising because, unlike NPIs, PPIs lack explicit licensors.
-Thus, in order to correctly predict the distribution of PPIs, the model would need to represent more complex aspects of language.
+Thus, in order to correctly predict the distribution of PPIs, the model would need to represent more complex aspects of language rather than solely establishing a relationship between \textit{not} and NPIs.
 
@@
 This suggests that, rather than representing the hierarchical structure of the sentence, the model approximates the NPI/licensor relation using a representation based on linear word order and precedence.
-Finally, in Section \ref{sec:intervening}, we saw that the model's ability to predict the distribution of NPIs was independent of the amount of intervening material.
-
-Despite the fact that the experiments in Section \ref{sec:c-command} suggested that the model does not represent the c-command relation, the results do not differ too dramatically from what is observed in psycholinguistic studies of NPI licensing.
-As we would expect, listeners with no time constraints tend to judge sentences like those in Examples \ref{ex:c-command-np} and \ref{ex:c-command-pp} as being ungrammatical.
-However, when fast time constraints are placed on the judgement task, some speakers mistakenly accept sentences like Example \ref{ex:c-command-np} (in which there exists a negative licensor, but it is contained in an embedded clause) as grammatical \cite{xiang2009illusory,xiang2013dependency,parker2016negative}.
-This suggests that, under time pressure, the human brain might process the NPI licensing relation using a heuristic similar to the one that our model has learned, which takes into account only precedence and not c-command.
+In Section \ref{sec:intervening}, we saw that the model's ability to predict the distribution of NPIs was independent of the amount of intervening material.
+Finally, in section \ref{sec:ppi} we observed that LSTMs do not assign a statistically significant difference in surprisal to positive polarity items in positive and negative contexts.
+This provides more evidence that the model merely understands that the words \textit{not} and \textit{no} often precede NPIs because PPIs are syntactically identical to NPIs, but they lack those explicit licensors.
 
 These results point towards a number of questions which could be addressed in future research.
-First of all, we might like to know how well these results generalize to other models.
+First of all, it could be interesting to compare our model's knowledge of the NPI/licensor relation with the knowledge of human subjects.
+Despite the fact that the experiments in Section \ref{sec:c-command} suggested that the model does not represent the c-command relation, the results do not differ too dramatically from what is observed in psycholinguistic studies of NPI licensing in humans.
+As we would expect, human subjects with no time constraints tend to judge sentences like those in Examples \ref{ex:c-command-np} and \ref{ex:c-command-pp} as being ungrammatical.
+However, when strict time constraints are placed on the judgement task, some speakers mistakenly accept sentences like Example \ref{ex:c-command-np} (in which there exists a negative licensor, but it is contained in an embedded clause) as grammatical \cite{xiang2009illusory,xiang2013dependency,parker2016negative}.
+These constructions which are mistakenly accepted as grammatical are referred to as Negative Polarity Illusions, and their existence suggests that, under time pressure, the human brain might process the NPI licensing relation using a heuristic similar to the one that our model has learned, which takes into account only precedence and not c-command.
+
+In order to analyse this apparent similarity more thoroughly, we could perform an experiment in which the same sentences are presented to both an LSTM language model and to a human subject.
+It has been observed that, in the general setting, surprisal in state of the art language models is correlated with human sentence processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+By focusing specifically on sentences like those in Example \ref{ex:c-command}, we could do a more controlled comparison of how LSTMs and human subjects respond to Negative Polarity Illusions.
+
+Another route for future research, which would not require the use of human test subjects, might involve studying how well our results generalize to other models.
 Our homegrown model was trained on a relatively small data set using limited resources; perhaps a larger LSTM like that presented in Jozefowicz et.
@@
 al.
-\shortcite{devlin2018bert} have both performed very well on a very general set of language modeling tasks.
-Repeating a similar analysis on GPT-2 or Bert could provide interesting insights into the extent to which transformer models represent hierarchical structure.
+\shortcite{devlin2018bert} have both performed very well on a very general set of language modeling tasks, and it has been suggested that they have some ability to represent hierarchical structure \cite{jawahar2019does,clark2019does}.
+Repeating a similar analysis on GPT-2 or Bert could provide interesting insights into the differences between LSTM models and transformer models, and the ability of these different model types to represent structural data.
 
